### 2024-04-18 paper/paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

\newpage

# References

### 2024-04-18 scripts/01-download_data.R
analysis_data added

@@ -1,21 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-library(nbastatR)
-library(tidyverse)
-
-#### Download data ####
-raw <- nba_players()
-
-#### Save data ####
-
-write_csv(raw, "inputs/data/player_data.csv") 
-
-         

### 2024-04-19 paper/paper.qmd
completed scripts

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Physical activity is as tied to obesity as we expect it to be"
+subtitle: "An analysis of American exerise habits and their BMI"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
-date: today
+  - Victor Ma
+thanks: "Code and data are available at: https://github.com/bestmustard/activity-bmi"
+date: 03/16/2024
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: ""
 format: pdf
@@
 bibliography: references.bib
+editor: 
+  markdown: 
+    wrap: 72
+
+header-includes:
+ \usepackage{float}
+ \floatplacement{figure}{H}
 ---
@@
 library(tidyverse)
-library(palmerpenguins)
+library(dataverse)
+library(knitr)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(dplyr)
+library(patchwork)
+library(ggplot2)
+library(broom.mixed)
+activity_data <- read_csv("../inputs/data/analysis_data/analysis_data.csv")
+activity_data$obese = as_factor(activity_data$obese)
+activity_data$activity = factor(
+      activity_data$activity,
+      levels = c("No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio")
+    )
 ```
@@
 
+## Data Preparation and Cleaning
+The 2020 CCES data was first obtained via the dataverse package directly from Harvard Dataverse, and then saved in parquet format using the arrow package for efficient storage and access [@arrow].
+The cleaning process involved filtering for registered voters who chose either Biden or Trump, in order to treat presidential votes as a binary outcome.
+
+The respondent variables (age, gender, and education) were transformed for clarity.
+Age was categorized into groups ("65+", "45-64", "30-44", "18-29"), gender was labeled ("Male", "Female"), and education levels were delineated ("No HS", "High school graduate", "Some college", "2-year", "4-year", "Post-grad").
+These steps utilized the dplyr package for data manipulation.
+
+The cleaned dataset was then saved in both CSV and parquet formats.
+
+The distributions for each explanatory variable are illustrated in @fig-activity and @fig-income below:
+
+```{r out.height="20%"}
+#| label: fig-activity
+#| fig-cap: "Number of respondents at each activity level"
+#| echo: false
+#| warning: false
+#| message: false
+
+ggplot(activity_data, aes(x = activity, fill = activity)) +
+  coord_flip()+
+  geom_bar() +
+  scale_fill_brewer(palette = "Paired") + # More colors for distinction
+  theme_minimal() +
+  labs(title = "Respondents per Level of Activity",
+       y = "Activity Level",
+       x = "Count") +
+  theme(legend.title = element_blank())
+
+```
+
+```{r out.height="20%"}
+#| label: fig-income
+#| fig-cap: "Number of respondents per income bracket"
+#| echo: false
+#| warning: false
+#| message: false
+ggplot(activity_data, aes(x = income)) +
+  geom_bar(fill = "#778899") +
+  theme_minimal() +
+  labs(title = "Respondents per income bracket", x = "Income (Thousands)", y = "Count")+
+  theme(legend.title = element_blank())
+
+```
+
+@fig-activity tells us that 4-year degree graduates are the most well represented in the polls, followed by high school graduates, people with some college, and then 2-year degree graduates.
+People who have not graduated high school are hardly represented at all.
+
+In @fig-income we see 45-64 being the most dominant age group, with much less representation in the 18-29 range. Also, about 20% more females participated in the study than males.
+
+Below are some figures representing who people voted for based on their age and gender, followed by education level and gender.
+
+```{r out.height="20%"}
+#| label: fig-activityobesity
+#| fig-cap: "Prevalence of obesity by activity level"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Relationship between Obesity and Activity Level
+ggplot(activity_data, aes(x = activity, fill = obese)) +
+  geom_bar(position = "fill") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  labs(title = "Obesity Proportion by Activity Levels", x = "Activity Level", y = "Proportion")
+```
+@fig-activityobesity shows a negative correlation between votes for Joe Biden and older age groups.
+```{r out.height="20%"}
+#| label: fig-incomeobesity
+#| fig-cap: "Prevalence of obesity by income level"
+#| echo: false
+#| warning: false
+#| message: false
+ggplot(activity_data, aes(x = income, fill = obese)) +
+  geom_bar(position = "fill") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  labs(title = "Obesity Proportion by Income Brackets", x = "Age Group", y = "Proportion")
+```
+We can see that in @fig-gendervotes, more women voted for Joe Biden than men.
+
+# Model
+
+Logistic regression is a model used when the outcome or dependent variable is binary, which fits this scenario perfectly as I am modelling the binary outcome of below 30 BMI or a BMI of 30 and above.
+
+The regression model will calculate the log odds of the probability that a person has a BMI of 30 or above, and then map it to a probability between 0 and 1 through the logistic function.
+
+The standard logistic function $\sigma(t)$ for a real-valued input $t$ is defined as:
+
+```{=tex}
+\[ \sigma(t) = \frac{1}{1 + e^{-t}} \]
+```
+
+The graph of the logistic function is an S-shaped curve known as a sigmoid curve. It approaches 1 as $t$ goes to positive infinity and approaches 0 as $t$ goes to negative infinity.
+
+In logistic regression, the input $t$ is the linear combination of predictors including the intercept, which can be represented as $\beta_0 + \beta_1X_1 + \beta_2X_2 + ...
++ \beta_nX_n$.
+The logistic function then translates this into a probability that the dependent variable is 1 (has a BMI above 30).
+
+In this situation, I will be using the predictors activity level and income and then applying the logistic function to get the probability $P(Y_i=1)$ that a respondent $i$ is considered obese.
+
+This model is particularly strong at handling categorical dependent variables, which each of my explanatory variables fall under [@Hosmer2013].
+
+## Model Specification
+
+The logistic regression model is defined as:
+
+```{=tex}
+\[
+\log\left(\frac{P(Y_i=1)}{1 - P(Y_i=1)}\right) = \beta_0 + \beta_1X_{\text{activity},i} + \beta_2X_{\text{income},i}
+\]
+```
+## Model set-up
+
+- $Y_i$ is the binary indicator of having a BMI of 30 or above (1) versus a BMI below 30 for respondent $i$.
+- $X_{\text{activity},i}$, $X_{\text{income},i}$ are the activity level and income of respondent $i$, respectively.
+- $\beta_0$ represents the model intercept, while $\beta_1$ and $\beta_2$ are coefficients quantifying the effects of activity level and income on the likelihood of being considered obese.
+
+I fit my logistic regression model to the data using 'stan_glm()' function from the 'rstanarm' package in R [@citeR, @Rstanarm].
+This function will automatically determine each of the $\beta$ coefficients in the model, using a smaller slice sample of 3000 from the data we processed.
+This function also uses Bayesian logistic regression with the default priors from 'rstanarm'.
+
+### Model Justification
+
+In order to better interpret the results of the model, I can create a coefficient plot to visually see the effect sizes of the predictor variables on the likelihood of an individual supporting Biden.
+
+@fig-coefficient maps each predictor variable on the y-axis to an effect size and confidence interval on the x-axis.
+The effect size is the change in log-odds of supporting Biden for a one-unit increase in the predictor variable, which is essentially how much impact each demographic characterstic has an effect on voting for Biden.
+
+The confidence intervals tell us the range within we can be confident that the true effect lies.
+Smaller confidence intervals means there is a higher level of precision in the estimate of the effect size.
+I am not as interested in the intervals that cross zero because that means that there is data to support each side (Biden or Trump) and so they are less statistically significant.
+
+```{r}
+#| label: fig-coefficient
+#| fig-cap: "Coefficient plot of demographics"
+#| echo: false
+#| warning: false
+#| message: false
+obesity_likelihood <- readRDS(file ="../models/obesity_likelihood.rds")
+
+# Assuming `tidy_obesity_likelihood` is a tidy version of your model with terms and estimate
+tidy_obesity_likelihood <- tidy(obesity_likelihood)
+
+ggplot(tidy_obesity_likelihood, aes(x = term, y = estimate)) +
+  geom_point(aes(color = estimate > 0), size = 4) +  # color points based on positive or negative coefficient
+  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +  #error bars
+  coord_flip() +  
+  scale_color_manual(values = c("blue", "red")) +  # Set colors for positive and negative coefficients
+  labs(title = "Obesity Likelihood",
+       x = "Predictor Variables",
+       y = "Effect on Log Odds of Being Considered Obese",
+       color = "Effect Direction") +
+  theme_minimal()  # Use a minimal theme
+
+# Ensure the 'tidy_obesity_likelihood' contains the columns you want to summarize
+tidy_obesity_likelihood <- broom::tidy(obesity_likelihood)
+
+# Check available columns
+names(tidy_obesity_likelihood)
+```
+
+The conclusions I can draw from @fig-coefficient align with what I expected given the dataset.
+It's clear that higher education levels, specifically 4-year degrees and post-grad degrees, are positively associated with support for Biden.
+Higher age groups, specifically 45-64 and 65+, are strongly tied to voting for Trump.
+Men are less likely to vote for Biden than Women.
+
+\newpage
+# Results
+
+Figures @fig-page, @fig-pgender, and @fig-pedu are recreations of the earlier graphs we saw showcasing the proportion of voters for Joe Biden by groups within each demographic, where the bar is the data from CCES 2020 while the point is the prediction generated by the model.
+
+```{r out.height="25%"}
+#| label: fig-page
+#| fig-cap: "Model prediction for votes by age vs. CCES 2020 data"
+#| echo: false
+#| warning: false
+#| message: false
+# Example constant values for the other factors
+constant_education <- "2-year" # Choose a representative value
+constant_gender <- "Male" # Choose a representative value
+
+# Create a data frame for age predictions
+age_data <- data.frame(age = c("18-29", "30-44", "45-64", "65+"),
+                       education = constant_education,
+                       gender = constant_gender)
+
+# Predict probabilities for each age group
+age_data$predicted_prob <- predict(obesity_likelihood, newdata = age_data, type = "response")
+library(ggplot2)
+# Actual proportions from the data
+activity_data_grouped <- activity_data %>%
+  group_by(age) %>%
+  summarize(Proportion_for_Biden = mean(voted_for == "Biden", na.rm = TRUE))
+
+# Assuming 'obesity_likelihood' is your logistic regression model object and
+# 'age_data' contains the predicted probabilities with 'age' as a factor
+age_data$age <- factor(age_data$age, levels = c("18-29", "30-44", "45-64", "65+"))
+
+# Create a combined plot
+ggplot() +
+  geom_bar(data = activity_data_grouped, aes(x = age, y = Proportion_for_Biden, fill = age), 
+           stat = "identity", position = "dodge", alpha = 0.5) +
+  geom_point(data = age_data, aes(x = age, y = predicted_prob, color = age), 
+             position = position_dodge(width = 0.9), size = 3) +
+  scale_fill_brewer(palette = "Set1") + 
+  scale_color_brewer(palette = "Set1") + 
+  labs(title = "Actual vs. Predicted Support for Biden by Age Group",
+       x = "Age Group", y = "Proportion") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+```{r out.height="25%"}
+#| label: fig-pgender
+#| fig-cap: "Model prediction for votes by gender vs. CCES2020 data"
+#| echo: false
+#| warning: false
+#| message: false
+# Example constant values for the other factors (choose representative values for each)
+constant_age <- "30-44" # Assuming this is a middle category for age
+
+# Create a data frame for gender predictions
+gender_data <- data.frame(
+  gender = c("Male", "Female"),
+  education = constant_education, # Assuming you have a constant value set for education
+  age = constant_age
+)
+
+# Predict probabilities for each gender category
+gender_data$predicted_prob <- predict(obesity_likelihood, newdata = gender_data, type = "response")
+
+# Create a data frame for education predictions
+education_data <- data.frame(
+  education = c("No HS", "High school graduate", "Some college", "2-year", "4-year", "Post-grad"),
+  gender = constant_gender, # Assuming you have a constant value set for gender
+  age = constant_age
+)
+
+# Predict probabilities for each education category
+education_data$predicted_prob <- predict(obesity_likelihood, newdata = education_data, type = "response")
+
+# Actual proportions by gender
+activity_data_gender_grouped <- activity_data %>%
+  group_by(gender) %>%
+  summarize(Proportion_for_Biden = mean(voted_for == "Biden", na.rm = TRUE))
+
+# Actual proportions by education
+activity_data_education_grouped <- activity_data %>%
+  group_by(education) %>%
+  summarize(Proportion_for_Biden = mean(voted_for == "Biden", na.rm = TRUE))
+# Plot for Gender
+ggplot() +
+  geom_bar(data = activity_data_gender_grouped, aes(x = gender, y = Proportion_for_Biden, fill = gender), 
+           stat = "identity", position = "dodge", alpha = 0.5) +
+  geom_point(data = gender_data, aes(x = gender, y = predicted_prob, color = gender), 
+             position = position_dodge(width = 0.9), size = 3) +
+  scale_fill_brewer(palette = "Set2") + 
+  scale_color_brewer(palette = "Set2") + 
+  labs(title = "Actual vs. Predicted Support for Biden by Gender",
+       x = "Gender", y = "Proportion") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+```{r out.height="25%"}
+#| label: fig-pedu
+#| fig-cap: "Model prediction for votes by education level vs. CCES2020 data"
+#| echo: false
+#| warning: false
+#| message: false
+# Plot for Education
+ggplot() +
+  geom_bar(data = activity_data_education_grouped, aes(x = education, y = Proportion_for_Biden, fill = education), 
+           stat = "identity", position = "dodge", alpha = 0.5) +
+  geom_point(data = education_data, aes(x = education, y = predicted_prob, color = education), 
+             position = position_dodge(width = 0.9), size = 3) +
+  scale_fill_brewer(palette = "Set3") + 
+  scale_color_brewer(palette = "Set3") + 
+  labs(title = "Actual vs. Predicted Support for Biden by Education",
+       x = "Education Level", y = "Proportion") +
+  theme_minimal() +
+  theme(legend.position = "none") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Adjust text angle for readability
+
+```
+
+By observation, it is evident that the model's prediction was fairly close to what the CCES data showed.
+The trend remains the same for each result.
+However, gender and education were predicted to have a higher sentiment of voting for Joe Biden while age had a lower sentiment.
+
+It is worth nothing though that in @fig-page, all age categories showed polling data of above 50% to vote for Biden.
+Now with the lower average vote, the older age categories of "45-64" and "65+" are leaning towards Trump.
+On @fig-pedu for the lower education groups, the "No-HS" group which was at just about 50% is now above it, and the "High School Graduate" is also closer to voting for Biden.
+While this may seem marginal, close elections where voter sentiment is close to 50% would be largely impacted by changes like this.
+Although ultimately this is not an accurate predictor for who would win the election, only which groups would vote for who.
+
+# Discussion
+
+This analysis demonstrates the demographics play in political preferences, with results that align with the existing literature.
+Our data suggests that age, education, and gender significantly influence electoral outcomes.
+In particular, the model's predictions closely mirrored CCES 2020 data trends, reinforcing the importance of these variables in understanding voter behavior.
+High education and younger ages showed high correlation with voting for joe biden.
+
+It is important to note that based on the sample size of our data (as seen in @fig-education and @fig-agegender), the majority age groups which participated in the survey were "45-64" and "65+", while the variables of education and gender were much more closer together.
+If this were representative of the actual population then it would suggest that the age groups are skewed more towards those that support Trump, so despite most of the proportions being towards Biden, it can be a lot more nuanced of there are more groups in demographics that support Trump.
+
+The demographic trends observed in our analysis, specifically the tendency of younger voters, women, and individuals with higher educational attainment towards supporting Joe Biden, mirror broader societal shifts.
+These findings align with historical patterns where demographic attributes have increasingly become predictors of political alignment.
+For instance, the growing political polarization along educational lines has been well-documented, with individuals possessing higher educational qualifications tending to support more progressive policies and candidates [@Dalton2020].
+
+Moreover, the gender gap in political preferences is not a novel phenomenon.
+Women have been consistently more likely to vote for Democratic candidates in recent elections, a trend that has been attributed to differing policy priorities, including healthcare, reproductive rights, and social welfare [@Schaffner2018].
+The 2020 election further highlighted this divide, as issues such as women's rights and gender equality took center stage, prompting a significant mobilization among female voters [@Coffe2020].
+
+Age demographics have also played a critical role in shaping the political landscape.
+The engagement and political leanings of younger voters have been subjects of extensive study, particularly given their inclination towards progressive and liberal policies.
+This has been partly driven by generational differences in value systems, with younger cohorts exhibiting greater concern for issues such as climate change, social justice, and economic inequality [@Norris2021].
+The 2020 election underscored the importance of the youth vote, which showed overwhelming support for Biden, reflecting a broader trend of political mobilization among younger Americans in response to current socio-political challenges [@Smith2020].
+
+The implications of these demographic trends extend beyond electoral outcomes.
+They signal shifting societal values and priorities, with potential long-term impacts on policy direction and political discourse.
+For instance, the increasing support for progressive policies among younger and more educated demographics may influence the policy agendas of political parties and candidates in future elections [@Inglehart2020].
+
+It's also crucial to note the limitations of relying solely on demographic variables to predict electoral outcomes.
+While age, education, and gender provide significant insights, the complex interplay of socio-economic factors, geographic location, and individual experiences also plays a crucial role in shaping political preferences.
+Future research should aim to incorporate a broader range of variables to capture the complexity of voter behavior more accurately.
+
+Our analysis contributes to the growing body of literature on demographic influences on political preferences, offering valuable insights into the dynamics of the 2020 U.S.
+presidential election.
+As the nation continues to navigate through periods of significant social and political change, understanding the demographic underpinnings of electoral outcomes will remain a critical area of study.
+
+# Weaknesses and Next Steps
+Although valuable, this study provides an intriguing analysis of how demographic influences affect political preferences; however, there are additional limitations to consider.
+Firstly, the study's analysis model, which used logistic regression, may have simplified the understanding of complex dynamics and could have overlooked some interactions between demographic variables that would provide a better understanding of voter behavior.
+Future research could utilize more advanced tools, such as multilevel or mixed-effects models, to more accurately capture the interplay of these demographic variables.
+
+Secondly, while detailed, the study's dataset may not fully account for the subtleties in voter behavior across all demographics.
+For example, the impact of regional urban-rural political contexts on voting patterns could have nuances beyond the current analysis's reach.
+Incorporating geographic and socioeconomic variables could further improve and strengthen the model, offering deeper insight into the electoral impact of demographic factors.
+
+Lastly, the study's focus on the 2020 election, though timely, limits its ability to capture longitudinal trends in voter behavior.
+A comparative analysis developed over multiple election cycles might reveal changes in demographic influences on political preferences over time, providing a richer context for interpreting the 2020 election results.
+In summary, this paper not only sheds light on the demographic determinants of voter preferences in a pivotal election year but also encourages reflection on the broader societal shifts these patterns indicate.
+As America faces deep social and political challenges, understanding the demographic underpinnings of political allegiance becomes increasingly important.
+Further research, building on the insights provided in this paper and addressing the weaknesses outlined, has the potential to offer even better explanations of the complex tapestry of American electoral politics.
 \newpage

### 2024-04-19 scripts/clean.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw 2020 CCES data recorded by Schaffner, Brian et.al 
 # Author: Victor Ma
-# Date: 15 Mar 2024
+# Date: 19 Apr 2024
 # Contact: victo.ma@mail.utoronto.ca
 
 #### Workspace setup ####
@@ -13,28 +13,47 @@ library(arrow)
 raw_data <-
   read_csv("inputs/data/raw_data/Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv")
 
+length(which(raw_data$Question == "Percent of adults who engage in no leisure-time physical activity"))
+
+
 cleaned_data <- 
   raw_data |> 
-  select ("Topic", "Question", "BMI") |>
+  select ("Topic", "Question", "BMI", "Income") |>
   filter(Topic == "Physical Activity - Behavior") |>
   mutate(
-    overweight = if_else(BMI >= 30, "Overweight", "Not Overweight"),
-    overweight = as_factor(overweight),
+    obese = if_else(BMI >= 30, "Yes", "No"),
+    obese = as_factor(obese),
     activity = case_when(
       Question == "Percent of adults who engage in no leisure-time physical activity" ~ "No Activity",
       Question == "Percent of adults who engage in muscle-strengthening activities on 2 or more days a week" ~ "Strength",
       Question == "Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)" ~ "Cardio",
       Question == "Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic physical activity and engage in muscle-strengthening activities on 2 or more days a week" ~ "Cardio + Strength",
-      Question == "Percent of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)" ~ "Heavy Cardio"
+      Question == "Percent of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)" ~ "Double Cardio"
     ), 
[…]

### 2024-04-19 scripts/model.R

@@ -1,11 +1,9 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Creates a logistic regression model
+# Author: Victor Ma
+# Date: 19 Apr 2024
+# Contact: victo.ma@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
@@ -13,25 +11,25 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-
+activity_data <- read_csv("inputs/data/analysis_data/analysis_data.csv")
+activity_data$obese <- as_factor(activity_data$obese)
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+set.seed(853)
+
+# Reducing the dataset for manageable computation
+activity_data_reduced <- activity_data |> 
+  slice_sample(n = 3000)
 
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+# Specifying the logistic regression model
+obesity_likelihood <- stan_glm(
+  obese ~ activity + income,
+  data = activity_data_reduced,
+  family = binomial(link = "logit"),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  seed = 853
 )
 
-
+# Saving and loading the model for future analysis
+saveRDS(obesity_likelihood, file = "models/obesity_likelihood.rds")

### 2024-04-19 paper/paper.qmd
finished scripts, updated paper, updated usage

--- 2024-04-18 paper/paper.qmd
@@
 # Introduction
+In the context of public health crises, the term 'pandemic' comes with the connotation of infectious diseases sweeping across global populations.
+Yet, the United States finds itself grappling with a pandemic of a different nature, with similarly far-reaching consequences: obesity.
+It is no secret that obesity comes with a multitude of direct health impacts- including increased risk of stroke, high blood pressure, type 2 diabetes, and even mental health problems like clinical depression and anxiety [@cdc].
+Obesity is recognized as a chronic complex disease defined by excessive fat deposits that can impair health.
+In particular, individuals with a Body Mass Index (BMI) of 30 or above are considered obese.
 
+America's problem with obesity is long-lasting, with a 30.5% rate of obesity observed in the early 2000s [@obesityfacts].
+Marketing companies have taken advantage of this disease through the spreading of false information, in the form of fad diets, devices and products tied to fat loss, and misleading nutritional claims that often prioritize profit over health.
+In the modern era, fitness knowledge has been democratized with the advent of social media and digital platforms aiding in dissemination of health information [@Johnson2019].
+Despite this, obesity in America has only gotten worse- increasing to 41.9% in 2020 as one of the top 10 most obese countries in the world [@cdc].
+
+Obesity is a multifaceted challenge with ties to lifestyle, socioeconomic factors, and access to health education and resources.
+In this study I use a logistic regression model in order to predict the odds that an individual is obese based on the amount of physical activity they do per week and their level of income.
+As logistic regression models are used to model binary outcomes, my outcome will be whether an individual has a BMI of 30 or above or not.
+
+The data I am using is from the Center of Disease Control and Prevention (CDC), a government-officiated service organization dedicated to health research in the United States.
+In particular, this dataset contains information on an adult's diet, physical activity, and weight status from CDC's Behavioral Risk Factor Surveillance System- America's premier system for collecting data about health-related risk behaviours conducted through telephone surveys [@dataset].
+I will be focusing on the variables of activity level and level of income as explanatory variables in the logistic regression model.
+The estimand will be the probability that an individual is obese based on these factors.
+
+My report is structured into four main sections following the introduction.
+In the first section, I describe the data utilized for my analysis, presenting the CDC dataset as well as graphs that show the distribution of the explanatory variables.
+The second section details the logistic regression model, including the rationale for its use and an interpretation of preliminary findings.
+Next I will analyse the variables' impact on obesity prevalence through the use of graphs and specific numerics from my results.
+Finally, I discuss the implications of my findings, address potential weaknesses in my study, and suggest directions for future research.
+
+This analysis is conducted using R, using several R packages to facilitate my analysis and presentation.
+This includes tidyverse for data manipulation and visualization, knitr for report generation, modelsummary for model interpretation, and rstanarm for Bayesian regression modeling [@citeR; @Dataverse; @Knitr; @Modelsummary; @Rstanarm; @Tidybayes; @Dplyr].
+Some portions including ggplot graphs and the "Discussion" section were written with the help of ChatGPT4 @chatgpt.
+\newpage
+
+# Data
+
+I used a dataset called "Nutrition, Physical Activity, and Obesity - Behavioral Risk Factor Surveillance System" pulled directly from the CDC website [@dataset].
+The .csv file obtained from the website contains 93250 data points with the relevant information of an individual's activity level based on their survey response, BMI, and various demographics.
+The dataset is owned by the Division of Nutrition, Physical Activity, Obesity (DNPAO), a division under the CDC which focuses directly on preventing chronic diseases by promoting better nutrition practices.
+
+## Source Reliability
+
+The Centers for Disease Control and Prevention (CDC) is recognized as a reliable source for obesity data for several reasons.
+The CDC's Division of Nutrition, Physical Activity, and Obesity conducts comprehensive surveillance and research to understand and address obesity, focusing on policy and environmental strategies to promote healthy eating and active living.
+The organization collects data from the largest scale health survey systems in the United States, including both the Behavioral Risk Factor Surveillance System (BRFSS) and the National Health and Nutrition Examination Survey (NHANES) [@cdcdata; @cdcinitiatives].
+
+The BRFSS conducts more than 400,000 interviews annually across the United States, offering state-level estimates of obesity prevalence among other health parameters.
+The large sample size it offers provides a better understanding of the population when compared to smaller samples [@brfss].
+
+## Limitations
+
+The CDC is a reputable government-associated organization but the data did not come without inherent limitations.
+As with any telephone survey, respondents are susceptible to lying which would represent false data points.
+In addition, even if respondents believe they are telling the truth, it is possible that they do not have an accurate measurement of, for example, their activity level.
+There is no information available about the  methods used to validate this information on the CDC website.
+
+The data used in this paper does not represent the full dataset, as data points had to be removed for any missing responses.
+The previously more robust dataset with 93250 data points was reduced to 10218 in this process.
+The categories available for both the respondent variables do not allow for some details, as the markers for physical activity were very specific.
+Many individuals may not adhere directly to the possible responses in the survey, and it is impossible to account for individuals who perform more physiccal activity than the provided options.
+The income levels also do not provide a broad perspective, with the maximum level being $75,000+.
+There is no evidence or rationale provided regarding why these options were chosen.
+
+## Variables of Interest
+
+### Activity Level 
+Physical activity is a determinant of energy expenditure and is fundamentally linked to obesity and weight management.
+Regular physical activity can significantly reduce the risk of becoming obese by increasing the number of calories the body uses for energy [@hillenergy].
+Conversely, sedentary lifestyles are closely associated with obesity due to low energy expenditure.
+Research has consistently shown that low physical activity levels are predictive of obesity development over time.
+Incorporating various forms of exercise, including strength training and aerobic activities, can aid in maintaining a healthy weight and preventing obesity [@sallisobesity].
+
+### Income
+Income level is a social determinant of health that influences obesity rates.
+Higher income levels often correlate with better access to healthy foods, recreational facilities, and health services, which can contribute to lower obesity rates [@pickett].
+Conversely, lower income levels are associated with limited access to healthy food options, reliance on cheaper, calorie-dense processed foods, and reduced opportunities for physical activity [@drewnowski].
+This economic disparity creates environments conducive to obesity development, particularly in communities where affordable healthy options are scarce.
+Research indicates that socioeconomic status, including income, plays a substantial role in the prevalence and distribution of obesity within populations.
+
+### Other Variables
+
+While I believed nutrition information would have been a suitable variable, the options provided in the dataset were only if the individual had "No Fruits" or "No Vegetables" in their regular diet.
+I did not think these two options were enough to make relevant conclusions.
+
+## Data Preparation and Cleaning
+The data was first downloaded as a .csv file directly from the CDC website and then saved in parquet format using the arrow package for efficient storage and access [@arrow].
+The cleaning process involved filtering for the columns for the relevant variables which were "Topic" ("Physical Activity - Behavior", "Obesity / Weight Status", "Fruits and Vegetables - Behavior"), "Question" (specific responses under the topic), BMI, and Income.
+
+The respondent variable of activity level was transformed for simplicity.
+Initially, the possible values included: "Percent of adults who engage in no leisure-time physical activity", "Percent of adults who engage in muscle-strengthening activities on 2 or more days a week", "Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)"...
+and so on.
+I removed the specifics and labelled them "No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio", as the markers of strength or cardio training remained the same throughout (strength meant 2 days of strength training, cardio meant 150 minutes of moderate intensity or 75 minutes of "vigorous-intensity").
+"Double Cardio" was named as such since it was defined as double the minutes of cardio as "Cardio".
+
+"NA" responses were then filtered out in order to make sure each data point contained all the variables used.
+
+The cleaned dataset was then saved in both CSV and parquet formats.
+
+The distributions for each explanatory variable are illustrated in @fig-activity and @fig-income below:
+
+```{r out.height="20%"}
+#| label: fig-activity
+#| fig-cap: "Number of respondents at each activity level"
+#| echo: false
+#| warning: false
+#| message: false
+
+ggplot(activity_data, aes(x = activity, fill = activity)) +
+  coord_flip()+
+  geom_bar() +
+  scale_fill_brewer(palette = "Paired") + # More colors for distinction
+  theme_minimal() +
+  labs(title = "Respondents per Level of Activity",
+       y = "Activity Level",
+       x = "Count") +
+  theme(legend.title = element_blank())
+
+```
+
+```{r out.height="20%"}
+#| label: fig-income
+#| fig-cap: "Number of respondents per income bracket"
+#| echo: false
+#| warning: false
+#| message: false
+ggplot(activity_data, aes(x = income)) +
+  geom_bar(fill = "#778899") +
+  theme_minimal() +
+  labs(title = "Respondents per income bracket", x = "Income (Thousands)", y = "Count")+
+  theme(legend.title = element_blank())
+
+```
+
+@fig-activity tells us that 4-year degree graduates are the most well represented in the polls, followed by high school graduates, people with some college, and then 2-year degree graduates.
+People who have not graduated high school are hardly represented at all.
+
+In @fig-income we see 45-64 being the most dominant age group, with much less representation in the 18-29 range. Also, about 20% more females participated in the study than males.
+
+Below are some figures representing who people voted for based on their age and gender, followed by education level and gender.
+
+```{r out.height="20%"}
+#| label: fig-activityobesity
+#| fig-cap: "Prevalence of obesity by activity level"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Relationship between Obesity and Activity Level
+ggplot(activity_data, aes(x = activity, fill = obese)) +
+  geom_bar(position = "fill") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  labs(title = "Obesity Proportion by Activity Levels", x = "Activity Level", y = "Proportion")
+```
+@fig-activityobesity shows a negative correlation between votes for Joe Biden and older age groups.
+```{r out.height="20%"}
+#| label: fig-incomeobesity
+#| fig-cap: "Prevalence of obesity by income level"
+#| echo: false
+#| warning: false
+#| message: false
+ggplot(activity_data, aes(x = income, fill = obese)) +
+  geom_bar(position = "fill") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  labs(title = "Obesity Proportion by Income Brackets", x = "Age Group", y = "Proportion")
+```
+We can see that in @fig-gendervotes, more women voted for Joe Biden than men.
+
+# Model
+
+Logistic regression is a model used when the outcome or dependent variable is binary, which fits this scenario perfectly as I am modelling the binary outcome of below 30 BMI or a BMI of 30 and above.
+
+The regression model will calculate the log odds of the probability that a person has a BMI of 30 or above, and then map it to a probability between 0 and 1 through the logistic function.
+
+The standard logistic function $\sigma(t)$ for a real-valued input $t$ is defined as:
+
+```{=tex}
+\[ \sigma(t) = \frac{1}{1 + e^{-t}} \]
+```
+
+The graph of the logistic function is an S-shaped curve known as a sigmoid curve. It approaches 1 as $t$ goes to positive infinity and approaches 0 as $t$ goes to negative infinity.
+
+In logistic regression, the input $t$ is the linear combination of predictors including the intercept, which can be represented as $\beta_0 + \beta_1X_1 + \beta_2X_2 + ...
++ \beta_nX_n$.
+The logistic function then translates this into a probability that the dependent variable is 1 (has a BMI above 30).
+
+In this situation, I will be using the predictors activity level and income and then applying the logistic function to get the probability $P(Y_i=1)$ that a respondent $i$ is considered obese.
+
+This model is particularly strong at handling categorical dependent variables, which each of my explanatory variables fall under [@Hosmer2013].
+
+## Model Specification
+
+The logistic regression model is defined as:
+
+```{=tex}
+\[
+\log\left(\frac{P(Y_i=1)}{1 - P(Y_i=1)}\right) = \beta_0 + \beta_1X_{\text{activity},i} + \beta_2X_{\text{income},i}
+\]
+```
+## Model set-up
+
+- $Y_i$ is the binary indicator of having a BMI of 30 or above (1) versus a BMI below 30 for respondent $i$.
+- $X_{\text{activity},i}$, $X_{\text{income},i}$ are the activity level and income of respondent $i$, respectively.
+- $\beta_0$ represents the model intercept, while $\beta_1$ and $\beta_2$ are coefficients quantifying the effects of activity level and income on the likelihood of being considered obese.
+
+I fit my logistic regression model to the data using 'stan_glm()' function from the 'rstanarm' package in R [@citeR, @Rstanarm].
+This function will automatically determine each of the $\beta$ coefficients in the model, using a smaller slice sample of 3000 from the data we processed.
+This function also uses Bayesian logistic regression with the default priors from 'rstanarm'.
+
+### Model Justification
+
+After creating the model, we can examine the summary in @fig-summary:
+
+```{r}
+#| label: fig-summary
+#| fig-cap: "Model summary"
+#| echo: false
+#| warning: false
+#| message: false
+obesity_likelihood <- readRDS(file ="../models/obesity_likelihood.rds")
+
+modelsummary(obesity_likelihood)
+```
+
+In order to better interpret the results of the model, I can create a coefficient plot to visually see the effect sizes of the predictor variables on the likelihood of an individual supporting Biden.
+
+@fig-coefficient maps each predictor variable on the y-axis to an effect size and confidence interval on the x-axis.
+The effect size is the change in log-odds of supporting Biden for a one-unit increase in the predictor variable, which is essentially how much impact each demographic characterstic has an effect on voting for Biden.
+
+The confidence intervals tell us the range within we can be confident that the true effect lies.
+Smaller confidence intervals means there is a higher level of precision in the estimate of the effect size.
+I am not as interested in the intervals that cross zero because that means that there is data to support each side (Biden or Trump) and so they are less statistically significant.
+
+```{r}
+#| label: fig-coefficient
+#| fig-cap: "Coefficient plot of demographics"
+#| echo: false
+#| warning: false
+#| message: false
+obesity_likelihood <- readRDS(file ="../models/obesity_likelihood.rds")
+
+# Assuming `tidy_obesity_likelihood` is a tidy version of your model with terms and estimate
+tidy_obesity_likelihood <- tidy(obesity_likelihood)
+
+ggplot(tidy_obesity_likelihood, aes(x = term, y = estimate)) +
+  geom_point(aes(color = estimate > 0), size = 4) +  # color points based on positive or negative coefficient
+  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +  #error bars
+  coord_flip() +  
+  scale_color_manual(values = c("green", "yellow")) +  # Set colors for positive and negative coefficients
+  labs(title = "Obesity Likelihood",
+       x = "Predictor Variables",
+       y = "Effect on Log Odds of Being Considered Obese",
+       color = "Effect Direction") +
+  theme_minimal()  # Use a minimal theme
+
+# Ensure the 'tidy_obesity_likelihood' contains the columns you want to summarize
+tidy_obesity_likelihood <- broom::tidy(obesity_likelihood)
+
+# Check available columns
+names(tidy_obesity_likelihood)
+```
+
+The conclusions I can draw from @fig-coefficient align with what I expected given the dataset.
+It's clear that higher education levels, specifically 4-year degrees and post-grad degrees, are positively associated with support for Biden.
+Higher age groups, specifically 45-64 and 65+, are strongly tied to voting for Trump.
+Men are less likely to vote for Biden than Women.
+
+\newpage
+# Results
+
+Figures @fig-page, @fig-pgender, and @fig-pedu are recreations of the earlier graphs we saw showcasing the proportion of voters for Joe Biden by groups within each demographic, where the bar is the data from CCES 2020 while the point is the prediction generated by the model.
+
+```{r out.height="25%"}
+#| label: fig-pactivity
+#| fig-cap: "Model prediction for obesity by income level vs. CDC data"
+#| echo: false
+#| warning: false
+#| message: false
+constant_activity <- "No Activity" # Choose a representative value
+
+income_data <- data.frame(income = c("0-15", "15-25", "25-35", "35-50", "50-75", "75+"),
+                       activity = constant_activity)
+
+income_data$predicted_prob <- predict(obesity_likelihood, newdata = income_data, type = "response")
+library(ggplot2)
+activity_data_grouped <- activity_data %>%
+  group_by(income) %>%
+  summarize(Obesity_Prediction = mean(obese == "Yes", na.rm = TRUE))
+
+income_data$income <- factor(income_data$income, levels = c("0-15", "15-25", "25-35", "35-50", "50-75", "75+"))
+
+ggplot() +
+  geom_bar(data = activity_data_grouped, aes(x = income, y = Obesity_Prediction, fill = income), 
+           stat = "identity", position = "dodge", alpha = 0.5) +
+  geom_point(data = income_data, aes(x = income, y = predicted_prob, color = income), 
+             position = position_dodge(width = 0.9), size = 3) +
+  scale_fill_brewer(palette = "Set1") + 
+  scale_color_brewer(palette = "Set1") + 
+  labs(title = "Actual vs. Predicted Obesity by Income Level",
+       x = "Income Level", y = "Proportion") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+```{r out.height="25%"}
+#| label: fig-pobesity
+#| fig-cap: "Model prediction for obesity by activity level vs. CDC data"
+#| echo: false
+#| warning: false
+#| message: false
+# Example constant values for the other factors (choose representative values for each)
+constant_income <- "35-50" # Choose a representative value
+
+activity_plot_data <- data.frame(activity = c("No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio"),
+                       income = constant_income)
+
+activity_plot_data$predicted_prob <- predict(obesity_likelihood, newdata = activity_plot_data, type = "response")
+library(ggplot2)
+activity_data_grouped <- activity_data %>%
+  group_by(activity) %>%
+  summarize(Obesity_Prediction = mean(obese == "Yes", na.rm = TRUE))
+
+activity_plot_data$activity <- factor(activity_plot_data$activity, levels = c("No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio"))
+
+ggplot() +
+  geom_bar(data = activity_data_grouped, aes(x = activity, y = Obesity_Prediction, fill = activity), 
+           stat = "identity", position = "dodge", alpha = 0.5) +
+  geom_point(data = activity_plot_data, aes(x = activity, y = predicted_prob, color = activity), 
+             position = position_dodge(width = 0.9), size = 3) +
+  scale_fill_brewer(palette = "Set1") + 
+  scale_color_brewer(palette = "Set1") + 
+  labs(title = "Actual vs. Predicted Obesity by Activity Level",
+       x = "Activity Level", y = "Proportion") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+By observation, it is evident that the model's prediction was fairly close to what the CCES data showed.
+The trend remains the same for each result.
+However, gender and education were predicted to have a higher sentiment of voting for Joe Biden while age had a lower sentiment.
+
+It is worth nothing though that in @fig-pactivity, all age categories showed polling data of above 50% to vote for Biden.
+Now with the lower average vote, the older age categories of "45-64" and "65+" are leaning towards Trump.
+On @fig-pobesity for the lower education groups, the "No-HS" group which was at just about 50% is now above it, and the "High School Graduate" is also closer to voting for Biden.
+While this may seem marginal, close elections where voter sentiment is close to 50% would be largely impacted by changes like this.
+Although ultimately this is not an accurate predictor for who would win the election, only which groups would vote for who.
+
+# Discussion
+
+# Weaknesses and Next Steps
 \newpage

### 2024-04-19 paper/paper.tex
finished scripts, updated paper, updated usage

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}

  \newcolumntype{d}{S[
    input-open-uncertainty=,
    input-close-uncertainty=,
    parse-numbers = false,
    table-align-text-pre=false,
    table-align-text-post=false
  ]}

\KOMAoption{captions}{tableheading}
\usepackage{float} \floatplacement{figure}{H}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Physical activity is as tied to obesity as we expect it to be},
  pdfauthor={Victor Ma},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Physical activity is as tied to obesity as we expect it to
be\thanks{Code and data are available at:
https://github.com/bestmustard/activity-bmi}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{An analysis of American exerise habits and their BMI}
\author{Victor Ma}
\date{March 16, 2024}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, enhanced, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In the context of public health crises, the term `pandemic' comes with
the connotation of infectious diseases sweeping across global
populations. Yet, the United States finds itself grappling with a
pandemic of a different nature, with similarly far-reaching
consequences: obesity. It is no secret that obesity comes with a
multitude of direct health impacts- including increased risk of stroke,
high blood pressure, type 2 diabetes, and even mental health problems
like clinical depression and anxiety ({``Health Effects of Overweight
and Obesity''} 2022). Obesity is recognized as a chronic complex disease
defined by excessive fat deposits that can impair health. In particular,
individuals with a Body Mass Index (BMI) of 30 or above are considered
obese.

America's problem with obesity is long-lasting, with a 30.5\% rate of
obesity observed in the early 2000s (Centers for Disease Control and
Prevention 2021). Marketing companies have taken advantage of this
disease through the spreading of false information, in the form of fad
diets, devices and products tied to fat loss, and misleading nutritional
claims that often prioritize profit over health. In the modern era,
fitness knowledge has been democratized with the advent of social media
and digital platforms aiding in dissemination of health information
(Johnson and Lee 2019). Despite this, obesity in America has only gotten
worse- increasing to 41.9\% in 2020 as one of the top 10 most obese
countries in the world ({``Health Effects of Overweight and Obesity''}
2022).

Obesity is a multifaceted challenge with ties to lifestyle,
socioeconomic factors, and access to health education and resources. In
this study I use a logistic regression model in order to predict the
odds that an individual is obese based on the amount of physical
activity they do per week and their level of income. As logistic
regression models are used to model binary outcomes, my outcome will be
whether an individual has a BMI of 30 or above or not.

The data I am using is from the Center of Disease Control and Prevention
(CDC), a government-officiated service organization dedicated to health
research in the United States. In particular, this dataset contains
information on an adult's diet, physical activity, and weight status
from CDC's Behavioral Risk Factor Surveillance System- America's premier
system for collecting data about health-related risk behaviours
conducted through telephone surveys (Centers for Disease Control and
Prevention (CDC), National Center for Chronic Disease Prevention and
Health Promotion, Division of Nutrition, Physical Activity, and Obesity
2023). I will be focusing on the variables of activity level and level
of income as explanatory variables in the logistic regression model. The
estimand will be the probability that an individual is obese based on
these factors.

My report is structured into four main sections following the
introduction. In the first section, I describe the data utilized for my
analysis, presenting the CDC dataset as well as graphs that show the
distribution of the explanatory variables. The second section details
the logistic regression model, including the rationale for its use and
an interpretation of preliminary findings. Next I will analyse the
variables' impact on obesity prevalence through the use of graphs and
specific numerics from my results. Finally, I discuss the implications
of my findings, address potential weaknesses in my study, and suggest
directions for future research.

This analysis is conducted using R, using several R packages to
facilitate my analysis and presentation. This includes tidyverse for
data manipulation and visualization, knitr for report generation,
modelsummary for model interpretation, and rstanarm for Bayesian
regression modeling (R Core Team 2021; \textbf{Dataverse?}; Xie 2021;
Arel-Bundock 2021; Goodrich et al. 2022; Kay 2021; Wickham et al. 2021).
Some portions including ggplot graphs and the ``Discussion'' section
were written with the help of ChatGPT4 OpenAI (2023). \newpage

\hypertarget{data}{%
\section{Data}\label{data}}

I used a dataset called ``Nutrition, Physical Activity, and Obesity -
Behavioral Risk Factor Surveillance System'' pulled directly from the
CDC website (Centers for Disease Control and Prevention (CDC), National
Center for Chronic Disease Prevention and Health Promotion, Division of
Nutrition, Physical Activity, and Obesity 2023). The .csv file obtained
from the website contains 93250 data points with the relevant
information of an individual's activity level based on their survey
response, BMI, and various demographics. The dataset is owned by the
Division of Nutrition, Physical Activity, Obesity (DNPAO), a division
under the CDC which focuses directly on preventing chronic diseases by
promoting better nutrition practices.

\hypertarget{source-reliability}{%
\subsection{Source Reliability}\label{source-reliability}}

The Centers for Disease Control and Prevention (CDC) is recognized as a
reliable source for obesity data for several reasons. The CDC's Division
of Nutrition, Physical Activity, and Obesity conducts comprehensive
surveillance and research to understand and address obesity, focusing on
policy and environmental strategies to promote healthy eating and active
living. The organization collects data from the largest scale health
survey systems in the United States, including both the Behavioral Risk
Factor Surveillance System (BRFSS) and the National Health and Nutrition
Examination Survey (NHANES) ({``Data \& Statistics \textbar{} Overweight
\& Obesity \textbar{} CDC,''} n.d.; {``Overweight \& Obesity \textbar{}
CDC,''} n.d.).

The BRFSS conducts more than 400,000 interviews annually across the
United States, offering state-level estimates of obesity prevalence
among other health parameters. The large sample size it offers provides
a better understanding of the population when compared to smaller
samples ({``4 Comparison of Data Sources Used to Assess Obesity
Prevalence and Trends,''} n.d.).

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

The CDC is a reputable government-associated organization but the data
did not come without inherent limitations. As with any telephone survey,
respondents are susceptible to lying which would represent false data
points. In addition, even if respondents believe they are telling the
truth, it is possible that they do not have an accurate measurement of,
for example, their activity level. There is no information available
about the methods used to validate this information on the CDC website.

The data used in this paper does not represent the full dataset, as data
points had to be removed for any missing responses. The previously more
robust dataset with 93250 data points was reduced to 10218 in this
process. The categories available for both the respondent variables do
not allow for some details, as the markers for physical activity were
very specific. Many individuals may not adhere directly to the possible
responses in the survey, and it is impossible to account for individuals
who perform more physiccal activity than the provided options. The
income levels also do not provide a broad perspective, with the maximum
level being \$75,000+. There is no evidence or rationale provided
regarding why these options were chosen.

\hypertarget{variables-of-interest}{%
\subsection{Variables of Interest}\label{variables-of-interest}}

\hypertarget{activity-level}{%
\subsubsection{Activity Level}\label{activity-level}}

Physical activity is a determinant of energy expenditure and is
fundamentally linked to obesity and weight management. Regular physical
activity can significantly reduce the risk of becoming obese by
increasing the number of calories the body uses for energy (Hill and
Peters 2003). Conversely, sedentary lifestyles are closely associated
with obesity due to low energy expenditure. Research has consistently
shown that low physical activity levels are predictive of obesity
development over time. Incorporating various forms of exercise,
including strength training and aerobic activities, can aid in
maintaining a healthy weight and preventing obesity (Sallis and Glanz
2012).

\hypertarget{income}{%
\subsubsection{Income}\label{income}}

Income level is a social determinant of health that influences obesity
rates. Higher income levels often correlate with better access to
healthy foods, recreational facilities, and health services, which can
contribute to lower obesity rates (Pickett and Wilkinson 2005).
Conversely, lower income levels are associated with limited access to
healthy food options, reliance on cheaper, calorie-dense processed
foods, and reduced opportunities for physical activity (Drewnowski and
Specter 2010). This economic disparity creates environments conducive to
obesity development, particularly in communities where affordable
healthy options are scarce. Research indicates that socioeconomic
status, including income, plays a substantial role in the prevalence and
distribution of obesity within populations.

\hypertarget{other-variables}{%
\subsubsection{Other Variables}\label{other-variables}}

While I believed nutrition information would have been a suitable
variable, the options provided in the dataset were only if the
individual had ``No Fruits'' or ``No Vegetables'' in their regular diet.
I did not think these two options were enough to make relevant
conclusions.

\hypertarget{data-preparation-and-cleaning}{%
\subsection{Data Preparation and
Cleaning}\label{data-preparation-and-cleaning}}

The data was first downloaded as a .csv file directly from the CDC
website and then saved in parquet format using the arrow package for
efficient storage and access (Richardson et al. 2024). The cleaning
process involved filtering for the columns for the relevant variables
which were ``Topic'' (``Physical Activity - Behavior'', ``Obesity /
Weight Status'', ``Fruits and Vegetables - Behavior''), ``Question''
(specific responses under the topic), BMI, and Income.

The respondent variable of activity level was transformed for
simplicity. Initially, the possible values included: ``Percent of adults
who engage in no leisure-time physical activity'', ``Percent of adults
who engage in muscle-strengthening activities on 2 or more days a
week'', ``Percent of adults who achieve at least 150 minutes a week of
moderate-intensity aerobic physical activity or 75 minutes a week of
vigorous-intensity aerobic activity (or an equivalent
combination)''\ldots{} and so on. I removed the specifics and labelled
them ``No Activity'', ``Strength'', ``Cardio'', ``Cardio + Strength'',
``Double Cardio'', as the markers of strength or cardio training
remained the same throughout (strength meant 2 days of strength
training, cardio meant 150 minutes of moderate intensity or 75 minutes
of ``vigorous-intensity''). ``Double Cardio'' was named as such since it
was defined as double the minutes of cardio as ``Cardio''.

``NA'' responses were then filtered out in order to make sure each data
point contained all the variables used.

The cleaned dataset was then saved in both CSV and parquet formats.

The distributions for each explanatory variable are illustrated in
Figure~\ref{fig-activity} and Figure~\ref{fig-income} below:

\begin{figure}

{\centering \includegraphics[width=\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-activity-1.pdf}

}

\caption{\label{fig-activity}Number of respondents at each activity
level}

\end{figure}

\begin{figure}

{\centering \includegraphics[width=\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-income-1.pdf}

}

\caption{\label{fig-income}Number of respondents per income bracket}

\end{figure}

Figure~\ref{fig-activity} tells us that 4-year degree graduates are the
most well represented in the polls, followed by high school graduates,
people with some college, and then 2-year degree graduates. People who
have not graduated high school are hardly represented at all.

In Figure~\ref{fig-income} we see 45-64 being the most dominant age
group, with much less representation in the 18-29 range. Also, about
20\% more females participated in the study than males.

Below are some figures representing who people voted for based on their
age and gender, followed by education level and gender.

\begin{figure}

{\centering \includegraphics[width=\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-activityobesity-1.pdf}

}

\caption{\label{fig-activityobesity}Prevalence of obesity by activity
level}

\end{figure}

Figure~\ref{fig-activityobesity} shows a negative correlation between
votes for Joe Biden and older age groups.

\begin{figure}

{\centering \includegraphics[width=\textwidth,height=0.2\textheight]{paper_files/figure-pdf/fig-incomeobesity-1.pdf}

}

\caption{\label{fig-incomeobesity}Prevalence of obesity by income level}

\end{figure}

We can see that in \textbf{?@fig-gendervotes}, more women voted for Joe
Biden than men.

\hypertarget{model}{%
\section{Model}\label{model}}

Logistic regression is a model used when the outcome or dependent
variable is binary, which fits this scenario perfectly as I am modelling
the binary outcome of below 30 BMI or a BMI of 30 and above.

The regression model will calculate the log odds of the probability that
a person has a BMI of 30 or above, and then map it to a probability
between 0 and 1 through the logistic function.

The standard logistic function \(\sigma(t)\) for a real-valued input
\(t\) is defined as:

\[ \sigma(t) = \frac{1}{1 + e^{-t}} \]

The graph of the logistic function is an S-shaped curve known as a
sigmoid curve. It approaches 1 as \(t\) goes to positive infinity and
approaches 0 as \(t\) goes to negative infinity.

In logistic regression, the input \(t\) is the linear combination of
predictors including the intercept, which can be represented as
\(\beta_0 + \beta_1X_1 + \beta_2X_2 + ... + \beta_nX_n\). The logistic
function then translates this into a probability that the dependent
variable is 1 (has a BMI above 30).

In this situation, I will be using the predictors activity level and
income and then applying the logistic function to get the probability
\(P(Y_i=1)\) that a respondent \(i\) is considered obese.

This model is particularly strong at handling categorical dependent
variables, which each of my explanatory variables fall under (Jr.,
Lemeshow, and Sturdivant 2013).

\hypertarget{model-specification}{%
\subsection{Model Specification}\label{model-specification}}

The logistic regression model is defined as:

\[
\log\left(\frac{P(Y_i=1)}{1 - P(Y_i=1)}\right) = \beta_0 + \beta_1X_{\text{activity},i} + \beta_2X_{\text{income},i}
\]

\hypertarget{model-set-up}{%
\subsection{Model set-up}\label{model-set-up}}

\begin{itemize}
\tightlist
\item
  \(Y_i\) is the binary indicator of having a BMI of 30 or above (1)
  versus a BMI below 30 for respondent \(i\).
\item
  \(X_{\text{activity},i}\), \(X_{\text{income},i}\) are the activity
  level and income of respondent \(i\), respectively.
\item
  \(\beta_0\) represents the model intercept, while \(\beta_1\) and
  \(\beta_2\) are coefficients quantifying the effects of activity level
  and income on the likelihood of being considered obese.
\end{itemize}

I fit my logistic regression model to the data using `stan\_glm()'
function from the `rstanarm' package in R Goodrich et al. (2022). This
function will automatically determine each of the \(\beta\) coefficients
in the model, using a smaller slice sample of 3000 from the data we
processed. This function also uses Bayesian logistic regression with the
default priors from `rstanarm'.

\hypertarget{model-justification}{%
\subsubsection{Model Justification}\label{model-justification}}

After creating the model, we can examine the summary in
Figure~\ref{fig-summary}:

\begin{figure}

{\centering

\hypertarget{fig-summary-1}{}
\begin{table}
\centering
\begin{tabular}[t]{lc}
\toprule
  & (1)\\
\midrule
(Intercept) & \num{0.060}\\
activityStrength & \num{0.176}\\
activityCardio & \num{5.373}\\
activityCardio + Strength & \num{-2.902}\\
activityDouble Cardio & \num{0.746}\\
income15-25 & \num{0.100}\\
income25-35 & \num{-0.717}\\
income35-50 & \num{-0.984}\\
income50-75 & \num{-1.058}\\
income75+ & \num{-0.396}\\
\midrule
Num.Obs. & \num{3000}\\
R2 & \num{0.340}\\
Log.Lik. & \num{-1435.801}\\
ELPD & \num{-1445.9}\\
ELPD s.e. & \num{25.8}\\
LOOIC & \num{2891.9}\\
LOOIC s.e. & \num{51.6}\\
WAIC & \num{2891.9}\\
RMSE & \num{0.40}\\
\bottomrule
\end{tabular}
\end{table}

}

\caption{\label{fig-summary}Model summary}

\end{figure}

In order to better interpret the results of the model, I can create a
coefficient plot to visually see the effect sizes of the predictor
variables on the likelihood of an individual supporting Biden.

Figure~\ref{fig-coefficient} maps each predictor variable on the y-axis
to an effect size and confidence interval on the x-axis. The effect size
is the change in log-odds of supporting Biden for a one-unit increase in
the predictor variable, which is essentially how much impact each
demographic characterstic has an effect on voting for Biden.

The confidence intervals tell us the range within we can be confident
that the true effect lies. Smaller confidence intervals means there is a
higher level of precision in the estimate of the effect size. I am not
as interested in the intervals that cross zero because that means that
there is data to support each side (Biden or Trump) and so they are less
statistically significant.

\begin{verbatim}
[1] "term"      "estimate"  "std.error"
\end{verbatim}

\begin{figure}

{\centering \includegraphics{paper_files/figure-pdf/fig-coefficient-1.pdf}

}

\caption{\label{fig-coefficient}Coefficient plot of demographics}

\end{figure}

The conclusions I can draw from Figure~\ref{fig-coefficient} align with
what I expected given the dataset. It's clear that higher education
levels, specifically 4-year degrees and post-grad degrees, are
positively associated with support for Biden. Higher age groups,
specifically 45-64 and 65+, are strongly tied to voting for Trump. Men
are less likely to vote for Biden than Women.

\newpage

\hypertarget{results}{%
\section{Results}\label{results}}

Figures \textbf{?@fig-page}, \textbf{?@fig-pgender}, and
\textbf{?@fig-pedu} are recreations of the earlier graphs we saw
showcasing the proportion of voters for Joe Biden by groups within each
demographic, where the bar is the data from CCES 2020 while the point is
the prediction generated by the model.

\begin{figure}

{\centering \includegraphics[width=\textwidth,height=0.25\textheight]{paper_files/figure-pdf/fig-pactivity-1.pdf}

}

\caption{\label{fig-pactivity}Model prediction for obesity by income
level vs.~CDC data}

\end{figure}

\begin{figure}

{\centering \includegraphics[width=\textwidth,height=0.25\textheight]{paper_files/figure-pdf/fig-pobesity-1.pdf}

}

\caption{\label{fig-pobesity}Model prediction for obesity by activity
level vs.~CDC data}

\end{figure}

By observation, it is evident that the model's prediction was fairly
close to what the CCES data showed. The trend remains the same for each
result. However, gender and education were predicted to have a higher
sentiment of voting for Joe Biden while age had a lower sentiment.

It is worth nothing though that in Figure~\ref{fig-pactivity}, all age
categories showed polling data of above 50\% to vote for Biden. Now with
the lower average vote, the older age categories of ``45-64'' and
``65+'' are leaning towards Trump. On Figure~\ref{fig-pobesity} for the
lower education groups, the ``No-HS'' group which was at just about 50\%
is now above it, and the ``High School Graduate'' is also closer to
voting for Biden. While this may seem marginal, close elections where
voter sentiment is close to 50\% would be largely impacted by changes
like this. Although ultimately this is not an accurate predictor for who
would win the election, only which groups would vote for who.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{weaknesses-and-next-steps}{%
\section{Weaknesses and Next Steps}\label{weaknesses-and-next-steps}}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-brfss}{}}%
{``4 Comparison of Data Sources Used to Assess Obesity Prevalence and
Trends.''} n.d. The National Academies of Sciences, Engineering,;
Medicine; \url{https://nap.nationalacademies.org/read/23505/chapter/7}.

\leavevmode\vadjust pre{\hypertarget{ref-Modelsummary}{}}%
Arel-Bundock, Vincent. 2021. \emph{Modelsummary: Summary Tables and
Plots for Statistical Models and Data: Beautiful, Customizable, and
Publication-Ready}.
\url{https://CRAN.R-project.org/package=modelsummary}.

\leavevmode\vadjust pre{\hypertarget{ref-obesityfacts}{}}%
Centers for Disease Control and Prevention. 2021. {``Adult Obesity
Facts.''}

\leavevmode\vadjust pre{\hypertarget{ref-dataset}{}}%
Centers for Disease Control and Prevention (CDC), National Center for
Chronic Disease Prevention and Health Promotion, Division of Nutrition,
Physical Activity, and Obesity. 2023. {``{Nutrition, Physical Activity,
and Obesity - Behavioral Risk Factor Surveillance System}.''} Atlanta,
Georgia:
\url{https://chronicdata.cdc.gov/Nutrition-Physical-Activity-and-Obesity/Nutrition-Physical-Activity-and-Obesity-Behavioral/hn4x-zwk7};
{Centers for Disease Control and Prevention}.

\leavevmode\vadjust pre{\hypertarget{ref-cdcdata}{}}%
{``Data \& Statistics \textbar{} Overweight \& Obesity \textbar{}
CDC.''} n.d. Centers for Disease Control; Prevention;
\url{https://www.cdc.gov/obesity/data/index.html}.

\leavevmode\vadjust pre{\hypertarget{ref-drewnowski}{}}%
Drewnowski, Adam, and S. E. Specter. 2010. {``Obesity and the Food
Environment: Dietary Energy Density and Diet Costs.''} \emph{American
Journal of Preventive Medicine} 27: 154--62.

\leavevmode\vadjust pre{\hypertarget{ref-Rstanarm}{}}%
Goodrich, Ben, Jonah Gabry, Imad Ali, and Sam Brilleman. 2022.
{``Rstanarm: {Bayesian} Applied Regression Modeling via {Stan}.''}
\url{https://mc-stan.org/rstanarm/}.

\leavevmode\vadjust pre{\hypertarget{ref-cdc}{}}%
{``Health Effects of Overweight and Obesity.''} 2022. Centers for
Disease Control; Prevention;
\url{https://www.cdc.gov/healthyweight/effects/index.html}.

\leavevmode\vadjust pre{\hypertarget{ref-hillenergy}{}}%
Hill, James O., and John C. Peters. 2003. {``Energy Balance and
Obesity.''} \emph{Circulation} 104: 51--52.

\leavevmode\vadjust pre{\hypertarget{ref-Johnson2019}{}}%
Johnson, Mark K., and Angela R. Lee. 2019. {``The Role of Digital Media
in Shaping Health Behaviors: Opportunities and Challenges.''}
\emph{Health Communication Today} 24 (5): 456--67.

\leavevmode\vadjust pre{\hypertarget{ref-Hosmer2013}{}}%
Jr., David W. Hosmer, Stanley Lemeshow, and Rodney X. Sturdivant. 2013.
\emph{Applied Logistic Regression}. 3rd ed. New York: John Wiley \&
Sons.

\leavevmode\vadjust pre{\hypertarget{ref-Tidybayes}{}}%
Kay, Matthew. 2021. \emph{Tidybayes: Tidy Data and Geoms for Bayesian
Models}. \url{https://CRAN.R-project.org/package=tidybayes}.

\leavevmode\vadjust pre{\hypertarget{ref-chatgpt}{}}%
OpenAI. 2023. {``ChatGPT: Optimizing Language Models for Dialogue.''}
\url{https://openai.com/}.

\leavevmode\vadjust pre{\hypertarget{ref-cdcinitiatives}{}}%
{``Overweight \& Obesity \textbar{} CDC.''} n.d. Centers for Disease
Control; Prevention; \url{https://www.cdc.gov/obesity/index.html}.

\leavevmode\vadjust pre{\hypertarget{ref-pickett}{}}%
Pickett, Kate E., and Richard G. Wilkinson. 2005. {``The Social
Determinants of Health: The Solid Facts.''} \emph{International Journal
of Epidemiology} 34: 1245.

\leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
R Core Team. 2021. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-arrow}{}}%
Richardson, Neal, Ian Cook, Nic Crane, Dewey Dunnington, Romain
François, Jonathan Keane, Dragoș Moldovan-Grünfeld, Jeroen Ooms, Jacob
Wujciak-Jens, and Apache Arrow. 2024. \emph{Arrow: Integration to
'Apache' 'Arrow'}. \url{https://github.com/apache/arrow/}.

\leavevmode\vadjust pre{\hypertarget{ref-sallisobesity}{}}%
Sallis, James F., and Karen Glanz. 2012. {``Role of Physical Activity in
the Prevention of Obesity in Children.''} \emph{International Journal of
Obesity} 14: 34--38.

\leavevmode\vadjust pre{\hypertarget{ref-Dplyr}{}}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2021.
\emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://CRAN.R-project.org/package=dplyr}.

\leavevmode\vadjust pre{\hypertarget{ref-Knitr}{}}%
Xie, Yihui. 2021. \emph{Knitr: A General-Purpose Package for Dynamic
Report Generation in r}. \url{https://CRAN.R-project.org/package=knitr}.

\end{CSLReferences}

\end{document}

### 2024-04-19 scripts/clean.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw 2020 CCES data recorded by Schaffner, Brian et.al 
+# Purpose: Cleans the physical activity and obesity data from CDC
 # Author: Victor Ma
 # Date: 19 Apr 2024
 # Contact: victo.ma@mail.utoronto.ca
@@ -13,8 +13,7 @@ library(arrow)
 raw_data <-
   read_csv("inputs/data/raw_data/Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv")
 
-length(which(raw_data$Question == "Percent of adults who engage in no leisure-time physical activity"))
-
+unique(raw_data$Topic)
 
 cleaned_data <- 
   raw_data |> 

### 2024-04-19 scripts/model.R

@@ -13,6 +13,10 @@ library(rstanarm)
 #### Read data ####
 activity_data <- read_csv("inputs/data/analysis_data/analysis_data.csv")
 activity_data$obese <- as_factor(activity_data$obese)
+activity_data$activity = factor(
+  activity_data$activity,
+  levels = c("No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio")
+)
 ### Model data ####
 set.seed(853)
 

### 2024-04-19 scripts/simulate.R

@@ -1,14 +1,51 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Simulates probability of being obese by activity level
+# Author: Victor Ma
+# Date: 19 April 2024
+# Contact: victo.ma@mail.utoronto.ca
+# In this simulation I use code from Alexander, Rohan 
 
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Simulate data ####
+set.seed(853)
+
+
+num_obs <- 1000 
+
+# Simulating data on obesity based on activity level and income
+simulated_obesity_data <- tibble(
+  income = sample(0:5, size = num_obs, replace = TRUE), # Simulate income categories
+  activity = sample(0:4, size = num_obs, replace = TRUE), # Simulate activity levels
+  obese_prob = ((income + activity) / 10) # Simplified probability model
+) %>%
+  mutate(
+    obese = if_else(runif(n = num_obs) < obese_prob, "Yes", "No"),
+    obese = as_factor(obese),
+    income = case_when(
+      income == 0 ~ "0-15",
+      income == 1 ~ "15-25",
+      income == 2 ~ "25-35",
+      income == 3 ~ "35-50",
+      income == 4 ~ "50-75",
+      income == 5 ~ "75+"
+    ),
+    activity = case_when(
+      activity == 0 ~ "No Activity",
+      activity == 1 ~ "Strength",
+      activity == 2 ~ "Cardio",
+      activity == 3 ~ "Cardio + Strength",
+      activity == 4 ~ "Double Cardio"
+    ),
+    income = factor(
+      income,
+      levels = c("0-15", "15-25", "25-35", "35-50", "50-75", "75+")
+    ),
+    activity = factor(
+      activity,
+      levels = c("No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio")
+    )
+  ) %>%
+  select(obese_prob, obese, activity, income)

### 2024-04-19 paper/paper.qmd
completed

@@
 ---
-title: "Physical activity is as tied to obesity as we expect it to be"
-subtitle: "An analysis of American exerise habits and their BMI"
+title: "Americans who are wealthier and perform more strength training are less likely to be obese"
+subtitle: "An analysis of how exercise habits and wealth correlate with BMI"
 author: 
@@
 date-format: long
-abstract: ""
+abstract: "This study examines the correlation between an individual's weekly physical activity, their income level, and their probability of being obese.
+The only correlation found between physical activity was with individuals who strength train having lower rates of obesity, and higher income levels tended to trend with lower obesity.
+The findings from this study can help researchers understand the mechanisms that drive weight loss, and also shows individuals that nutrition is a greater factor than exercise with respect to body composition."
 format: pdf
@@
 library(tidyverse)
-library(dataverse)
 library(knitr)
@@
 This analysis is conducted using R, using several R packages to facilitate my analysis and presentation.
-This includes tidyverse for data manipulation and visualization, knitr for report generation, modelsummary for model interpretation, and rstanarm for Bayesian regression modeling [@citeR; @Dataverse; @Knitr; @Modelsummary; @Rstanarm; @Tidybayes; @Dplyr].
-Some portions including ggplot graphs and the "Discussion" section were written with the help of ChatGPT4 @chatgpt.
+This includes tidyverse for data manipulation and visualization, knitr for report generation, modelsummary for model interpretation, and rstanarm for Bayesian regression modeling [@citeR; @Knitr; @Modelsummary; @Rstanarm; @Tidybayes; @Dplyr].
+Some portions including ggplot graphs and the "Data" and "Discussion" sections were written with the help of ChatGPT4 @chatgpt.
+
 \newpage
@@
 
-## Source Reliability
-
-The Centers for Disease Control and Prevention (CDC) is recognized as a reliable source for obesity data for several reasons.
 The CDC's Division of Nutrition, Physical Activity, and Obesity conducts comprehensive surveillance and research to understand and address obesity, focusing on policy and environmental strategies to promote healthy eating and active living.
-The organization collects data from the largest scale health survey systems in the United States, including both the Behavioral Risk Factor Surveillance System (BRFSS) and the National Health and Nutrition Examination Survey (NHANES) [@cdcdata; @cdcinitiatives].
-
-The BRFSS conducts more than 400,000 interviews annually across the United States, offering state-level estimates of obesity prevalence among other health parameters.
-The large sample size it offers provides a better understanding of the population when compared to smaller samples [@brfss].
+The organization collects data from the largest scale health survey systems in the United States, including both the Behavioral Risk Factor Surveillance System (BRFSS) and the National Health and Nutrition Examination Survey (NHANES) [@cdcdata].
 
@@
 The categories available for both the respondent variables do not allow for some details, as the markers for physical activity were very specific.
-Many individuals may not adhere directly to the possible responses in the survey, and it is impossible to account for individuals who perform more physiccal activity than the provided options.
+Many individuals may not adhere directly to the possible responses in the survey, and it is impossible to account for individuals who perform more physical activity than the provided options.
 The income levels also do not provide a broad perspective, with the maximum level being $75,000+.
@@
 
-@fig-activity tells us that 4-year degree graduates are the most well represented in the polls, followed by high school graduates, people with some college, and then 2-year degree graduates.
-People who have not graduated high school are hardly represented at all.
+@fig-activity tells us that Americans who do not perform any leisurely exercise are the most well represented, with all other groups having an equal number of points in this data.
 
-In @fig-income we see 45-64 being the most dominant age group, with much less representation in the 18-29 range. Also, about 20% more females participated in the study than males.
+In @fig-income, oddly enough every income level was equally represented in this study.
 
-Below are some figures representing who people voted for based on their age and gender, followed by education level and gender.
+Below are some figures representing the proportion of people who were considered obese based on their activity level and their wealth.
 
@@
 ```
-@fig-activityobesity shows a negative correlation between votes for Joe Biden and older age groups.
+@fig-activityobesity shows no correlation between obesity and activity level. The proportion of people with obesity doing moderate cardio is unexpected.
+
 ```{r out.height="20%"}
@@
 ```
-We can see that in @fig-gendervotes, more women voted for Joe Biden than men.
+We can see that in @fig-incomeobesity, obesity tended to trend down with higher levels of income.
+\newpage
 
@@
 ```
+
 ## Model set-up
@@
 
-```{r}
+```{r out.height="20%"}
 #| label: fig-summary
@@
 
-In order to better interpret the results of the model, I can create a coefficient plot to visually see the effect sizes of the predictor variables on the likelihood of an individual supporting Biden.
+#### Summary Interpretation
+
+Each coefficient in a logistic regression model quantifies the change in the log odds of the outcome per unit change in the predictor.
+In this context, with the probability of being obese as the outcome:
+
+(Intercept) (5.329): When all other variables are at their baseline levels (no physical activity and the lowest income category), the log odds of being obese are 5.329.
+
+activityCardio + Strength (-8.164): Engaging in both cardio and strength activities is associated with an 8.164 unit decrease in the log odds of being obese compared to the baseline of no activity.
+
+activityDouble Cardio (-4.518): Doing 300 minutes of moderate cardio or 150 minutes of vigorous cardio results in a 4.518 unit decrease in the log odds of being obese compared to the baseline.
+
+activityNo Activity (-5.269): Having no leisure-time physical activity is associated with a 5.269 unit decrease in the log odds of being obese compared to the baseline.
+This negative value is unexpected, as we would typically predict no activity to increase the odds of obesity unless the baseline includes negative health behaviors even more associated with obesity.
+
+activityStrength (-5.092): Engaging only in strength training activities is associated with a 5.092 unit decrease in the log odds of being obese compared to the baseline.
+
+income15-25 (0.096): Being in the \$15,000 to $25,000 income bracket is associated with a 0.096 unit increase in the log odds of being obese compared to the reference income category.
+
+income25-35 (-0.719): Being in the \$25,000 to $35,000 income bracket is associated with a 0.719 unit decrease in the log odds of being obese.
+
+income35-50 (-0.990): Being in the \$35,000 to $50,000 income bracket is associated with a 0.990 unit decrease in the log odds of being obese.
+
+income50-75 (-1.063): Being in the \$50,000 to $75,000 income bracket is associated with a 1.063 unit decrease in the log odds of being obese.
+
+income75+ (-0.401): Being in the over $75,000 income bracket is associated with a 0.401 unit decrease in the log odds of being obese.
+
+It is important to note that these are all associative relationships and don't imply causation.
+
+In order to better interpret the results of the model, I can create a coefficient plot to visually see the effect sizes of the predictor variables on the likelihood of an individual being obese.
 
 @fig-coefficient maps each predictor variable on the y-axis to an effect size and confidence interval on the x-axis.
-The effect size is the change in log-odds of supporting Biden for a one-unit increase in the predictor variable, which is essentially how much impact each demographic characterstic has an effect on voting for Biden.
+The effect size is the change in log-odds of being obese for a one-unit increase in the predictor variable, which is essentially how much impact each variable has an effect on being obese
 
@@
 Smaller confidence intervals means there is a higher level of precision in the estimate of the effect size.
-I am not as interested in the intervals that cross zero because that means that there is data to support each side (Biden or Trump) and so they are less statistically significant.
+I am not as interested in the intervals that cross zero because that means that there is data to support each side (obese and not obese) and so they are less statistically significant.
 
@@
 The conclusions I can draw from @fig-coefficient align with what I expected given the dataset.
-It's clear that higher education levels, specifically 4-year degrees and post-grad degrees, are positively associated with support for Biden.
-Higher age groups, specifically 45-64 and 65+, are strongly tied to voting for Trump.
-Men are less likely to vote for Biden than Women.
+We can see that generally, higher incomes trend with a lower log-likelihood of being obese.
+The conclusions drawn from the activity levels also are in line with what we saw earlier on @fig-activity, though they are not at all what I expected as I believed exercise to have a negative correlation with obesity.
+I did not expect that the (Intercept) no
 
@@
 
-Figures @fig-page, @fig-pgender, and @fig-pedu are recreations of the earlier graphs we saw showcasing the proportion of voters for Joe Biden by groups within each demographic, where the bar is the data from CCES 2020 while the point is the prediction generated by the model.
+Figures @fig-pactivity and @fig-pobesity, are recreations of the earlier graphs we saw showcasing the proportion of obesity by groups within categories, where the bar is the data from CDC while the point is the prediction generated by the model.
 
@@
 
-By observation, it is evident that the model's prediction was fairly close to what the CCES data showed.
-The trend remains the same for each result.
-However, gender and education were predicted to have a higher sentiment of voting for Joe Biden while age had a lower sentiment.
-
-It is worth nothing though that in @fig-pactivity, all age categories showed polling data of above 50% to vote for Biden.
-Now with the lower average vote, the older age categories of "45-64" and "65+" are leaning towards Trump.
-On @fig-pobesity for the lower education groups, the "No-HS" group which was at just about 50% is now above it, and the "High School Graduate" is also closer to voting for Biden.
-While this may seem marginal, close elections where voter sentiment is close to 50% would be largely impacted by changes like this.
-Although ultimately this is not an accurate predictor for who would win the election, only which groups would vote for who.
+We can see that the model prediction resembled the CDC data by trend, however for most of the categories within both @fig-pactivity and @fig-pobesity, the prediction for obesity was lower than the data suggested.
+Moderate cardio having such a high proportion of obesity as unexpected.
 
@@
 
+In this paper, we used real-world statistics on Americans' activity levels, income, and whether or not they were considered obese to find a correlation between these behavioural and socioeconomic factors and obesity.
+What we found was that contrary to popular belief, our model did not predict lower obesity levels for individuals who exercised more.
+In particular, there was a large proportion of individuals who did moderate cardio and were obese, but there was correlation with lower obesity and strength training.
+Generally, higher income levels were also associated with lower rates of obesity.
+
+While this data may not be indicative of reality, the lack of correlation between exercise and obesity may reveal truths about societal beliefs in terms of how to get fit.
+Like myself, many Americans grew up watching TV where commercial ad breaks would feature fad diets, fast 10 minute work outs, and buzz phrases like "do this for 30 minutes a week to lose X lbs of fat!" Fitness may be viewed with a dogmatic stigma due to media portrayals which push the idea of a 'hardcore' lifestyle being required for a fit body.
+
+Due to this, it is possible that individuals believe in exercise being a greater factor for fat loss than nutrition.
+With the highest category of 300 minutes of moderate cardio used in CDC's survey, an adult weighing 160 lbs can expect to burn approximately 1825 calories in a week [@mayoclinic].
+As a single pound of body fat contains roughly 3500 calories, this would seem like half a pound of fat loss per week.
+However, this does not account for the possibility that an individual consumes more food with added physical activity, and 1825 calories spread out over a week is obly 260 calories per day.
+That equates to less than a small fries at McDonalds, or 2 or 3 eggs with oil.
+
+It is impossible to predict how an individual's body composition will change without knowing their nutritional information, specifically their caloric intake.
+There are various reasons why higher incomes might trend with lower obesity rates.
+People with higher incomes also tend to have achieved higher levels of education, which may be tied to access of information.
+They also have more access to help from professionals, better training facilities, and more food options.
+The study "Nutrition quality of food purchases varies by household income: the SHoPPER study" published in BMC Public Health highlights that lower-income households tend to purchase foods of lower nutritional quality compared to higher-income households.
+This is due to financial constraints limiting the purchase of healthier options like fruits and vegetables, leading to a higher purchase of less healthful foods such as frozen desserts or fast food.
+The study emphasizes that food purchasing patterns significantly mediate income differences in dietary intake quality [@nutritionquality].
+
 # Weaknesses and Next Steps
+As outlined previously, the data points were self reported which could lead to false data due to dishonesty or lack of care in measurement.
+The dataset was also filtered to remove incomplete datapoints, resulting in almost 90% of the initial dataset being removed.
+The discrete categories used for both physical activity and income may oversimplify the spectrum of exercise habits which overlooks the nuances of individual physical activity patterns.
+The income cap at $75,000 also fails to show variations in prevalence of obesity in higher income brackets.
+
+Future research should aim to incorporate objective measures of physical activity, perhaps through wearable technology, to diminish self-reporting biases.
+A broader dataset, possibly integrating direct measures of physical activity and detailed nutritional intake, would enable a more comprehensive analysis.
+Investigating the impact of higher income brackets beyond $75,000 and accounting for regional cost-of-living differences could refine understanding of the income-obesity relationship.
+Additionally, longitudinal data could shed light on the temporal dynamics between exercise habits and BMI changes over time.
+
+Further, integrating geospatial analyses to assess environmental factors, such as access to recreational spaces and healthy food outlets, could add further context to obesity determinants.
+Given the multifaceted nature of obesity, interdisciplinary studies combining data from healthcare, urban planning, and social sciences could offer more holistic insights.
 \newpage

### 2024-04-19 paper/paper.tex
completed

@@
 \hypersetup{
-  pdftitle={Physical activity is as tied to obesity as we expect it to be},
+  pdftitle={Americans who are wealthier and perform more strength training are less likely to be obese},
   pdfauthor={Victor Ma},
@@
 
-\title{Physical activity is as tied to obesity as we expect it to
-be\thanks{Code and data are available at:
+\title{Americans who are wealthier and perform more strength training
+are less likely to be obese\thanks{Code and data are available at:
 https://github.com/bestmustard/activity-bmi}}
@@
 \makeatother
-\subtitle{An analysis of American exerise habits and their BMI}
+\subtitle{An analysis of how exercise habits and wealth correlate with
+BMI}
 \author{Victor Ma}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, enhanced, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
+\begin{abstract}
+This study examines the correlation between varying levels of physical
+activity,
+\end{abstract}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, boxrule=0pt, frame hidden, sharp corners, enhanced, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 conducted through telephone surveys (Centers for Disease Control and
-Prevention (CDC), National Center for Chronic Disease Prevention and
-Health Promotion, Division of Nutrition, Physical Activity, and Obesity
-2023). I will be focusing on the variables of activity level and level
-of income as explanatory variables in the logistic regression model. The
-estimand will be the probability that an individual is obese based on
-these factors.
+Prevention (CDC) 2023). I will be focusing on the variables of activity
+level and level of income as explanatory variables in the logistic
+regression model. The estimand will be the probability that an
+individual is obese based on these factors.
 
@@
 modelsummary for model interpretation, and rstanarm for Bayesian
-regression modeling (R Core Team 2021; \textbf{Dataverse?}; Xie 2021;
-Arel-Bundock 2021; Goodrich et al. 2022; Kay 2021; Wickham et al. 2021).
-Some portions including ggplot graphs and the ``Discussion'' section
-were written with the help of ChatGPT4 OpenAI (2023). \newpage
+regression modeling (R Core Team 2021; Xie 2021; Arel-Bundock 2021;
+Goodrich et al. 2022; Kay 2021; Wickham et al. 2021). Some portions
+including ggplot graphs and the ``Data'' and ``Discussion'' sections
+were written with the help of ChatGPT4 OpenAI (2023).
+
+\newpage
 
@@
 Behavioral Risk Factor Surveillance System'' pulled directly from the
-CDC website (Centers for Disease Control and Prevention (CDC), National
-Center for Chronic Disease Prevention and Health Promotion, Division of
-Nutrition, Physical Activity, and Obesity 2023). The .csv file obtained
-from the website contains 93250 data points with the relevant
-information of an individual's activity level based on their survey
-response, BMI, and various demographics. The dataset is owned by the
-Division of Nutrition, Physical Activity, Obesity (DNPAO), a division
-under the CDC which focuses directly on preventing chronic diseases by
-promoting better nutrition practices.
+CDC website (Centers for Disease Control and Prevention (CDC) 2023). The
+.csv file obtained from the website contains 93250 data points with the
+relevant information of an individual's activity level based on their
+survey response, BMI, and various demographics. The dataset is owned by
+the Division of Nutrition, Physical Activity, Obesity (DNPAO), a
+division under the CDC which focuses directly on preventing chronic
+diseases by promoting better nutrition practices.
 
-\hypertarget{source-reliability}{%
-\subsection{Source Reliability}\label{source-reliability}}
-
-The Centers for Disease Control and Prevention (CDC) is recognized as a
-reliable source for obesity data for several reasons. The CDC's Division
-of Nutrition, Physical Activity, and Obesity conducts comprehensive
-surveillance and research to understand and address obesity, focusing on
-policy and environmental strategies to promote healthy eating and active
-living. The organization collects data from the largest scale health
-survey systems in the United States, including both the Behavioral Risk
-Factor Surveillance System (BRFSS) and the National Health and Nutrition
-Examination Survey (NHANES) ({``Data \& Statistics \textbar{} Overweight
-\& Obesity \textbar{} CDC,''} n.d.; {``Overweight \& Obesity \textbar{}
-CDC,''} n.d.).
-
-The BRFSS conducts more than 400,000 interviews annually across the
-United States, offering state-level estimates of obesity prevalence
-among other health parameters. The large sample size it offers provides
-a better understanding of the population when compared to smaller
-samples ({``4 Comparison of Data Sources Used to Assess Obesity
-Prevalence and Trends,''} n.d.).
+The CDC's Division of Nutrition, Physical Activity, and Obesity conducts
+comprehensive surveillance and research to understand and address
+obesity, focusing on policy and environmental strategies to promote
+healthy eating and active living. The organization collects data from
+the largest scale health survey systems in the United States, including
+both the Behavioral Risk Factor Surveillance System (BRFSS) and the
+National Health and Nutrition Examination Survey (NHANES) ({``Data \&
+Statistics \textbar{} Overweight \& Obesity \textbar{} CDC,''} n.d.).
 
@@
 responses in the survey, and it is impossible to account for individuals
-who perform more physiccal activity than the provided options. The
-income levels also do not provide a broad perspective, with the maximum
-level being \$75,000+. There is no evidence or rationale provided
-regarding why these options were chosen.
+who perform more physical activity than the provided options. The income
+levels also do not provide a broad perspective, with the maximum level
+being \$75,000+. There is no evidence or rationale provided regarding
+why these options were chosen.
 
@@
 
-Figure~\ref{fig-activity} tells us that 4-year degree graduates are the
-most well represented in the polls, followed by high school graduates,
-people with some college, and then 2-year degree graduates. People who
-have not graduated high school are hardly represented at all.
+Figure~\ref{fig-activity} tells us that Americans who do not perform any
+leisurely exercise are the most well represented, with all other groups
+having an equal number of points in this data.
 
-In Figure~\ref{fig-income} we see 45-64 being the most dominant age
-group, with much less representation in the 18-29 range. Also, about
-20\% more females participated in the study than males.
+In Figure~\ref{fig-income}, oddly enough every income level was equally
+represented in this study.
 
-Below are some figures representing who people voted for based on their
-age and gender, followed by education level and gender.
+Below are some figures representing the proportion of people who were
+considered obese based on their activity level and their wealth.
 
@@
 
-Figure~\ref{fig-activityobesity} shows a negative correlation between
-votes for Joe Biden and older age groups.
+Figure~\ref{fig-activityobesity} shows no correlation between obesity
+and activity level. The proportion of people with obesity doing moderate
+cardio is unexpected.
 
@@
 
-We can see that in \textbf{?@fig-gendervotes}, more women voted for Joe
-Biden than men.
+We can see that in Figure~\ref{fig-incomeobesity}, obesity tended to
+trend down with higher levels of income. \newpage
 
@@
 
+\hypertarget{summary-interpretation}{%
+\paragraph{Summary Interpretation}\label{summary-interpretation}}
+
+Each coefficient in a logistic regression model quantifies the change in
+the log odds of the outcome per unit change in the predictor. In this
+context, with the probability of being obese as the outcome:
+
+(Intercept) (5.329): When all other variables are at their baseline
+levels (no physical activity and the lowest income category), the log
+odds of being obese are 5.329.
+
+activityCardio + Strength (-8.164): Engaging in both cardio and strength
+activities is associated with an 8.164 unit decrease in the log odds of
+being obese compared to the baseline of no activity.
+
+activityDouble Cardio (-4.518): Doing 300 minutes of moderate cardio or
+150 minutes of vigorous cardio results in a 4.518 unit decrease in the
+log odds of being obese compared to the baseline.
+
+activityNo Activity (-5.269): Having no leisure-time physical activity
+is associated with a 5.269 unit decrease in the log odds of being obese
+compared to the baseline. This negative value is unexpected, as we would
+typically predict no activity to increase the odds of obesity unless the
+baseline includes negative health behaviors even more associated with
+obesity.
+
+activityStrength (-5.092): Engaging only in strength training activities
+is associated with a 5.092 unit decrease in the log odds of being obese
+compared to the baseline.
+
+income15-25 (0.096): Being in the \$15,000 to \$25,000 income bracket is
+associated with a 0.096 unit increase in the log odds of being obese
+compared to the reference income category.
+
+income25-35 (-0.719): Being in the \$25,000 to \$35,000 income bracket
+is associated with a 0.719 unit decrease in the log odds of being obese.
+
+income35-50 (-0.990): Being in the \$35,000 to \$50,000 income bracket
+is associated with a 0.990 unit decrease in the log odds of being obese.
+
+income50-75 (-1.063): Being in the \$50,000 to \$75,000 income bracket
+is associated with a 1.063 unit decrease in the log odds of being obese.
+
+income75+ (-0.401): Being in the over \$75,000 income bracket is
+associated with a 0.401 unit decrease in the log odds of being obese.
+
+It is important to note that these are all associative relationships and
+don't imply causation.
+
 In order to better interpret the results of the model, I can create a
 coefficient plot to visually see the effect sizes of the predictor
-variables on the likelihood of an individual supporting Biden.
+variables on the likelihood of an individual being obese.
 
@@
 to an effect size and confidence interval on the x-axis. The effect size
-is the change in log-odds of supporting Biden for a one-unit increase in
-the predictor variable, which is essentially how much impact each
-demographic characterstic has an effect on voting for Biden.
+is the change in log-odds of being obese for a one-unit increase in the
+predictor variable, which is essentially how much impact each variable
+has an effect on being obese
 
@@
 as interested in the intervals that cross zero because that means that
-there is data to support each side (Biden or Trump) and so they are less
-statistically significant.
+there is data to support each side (obese and not obese) and so they are
+less statistically significant.
 
@@
 The conclusions I can draw from Figure~\ref{fig-coefficient} align with
-what I expected given the dataset. It's clear that higher education
-levels, specifically 4-year degrees and post-grad degrees, are
-positively associated with support for Biden. Higher age groups,
-specifically 45-64 and 65+, are strongly tied to voting for Trump. Men
-are less likely to vote for Biden than Women.
+what I expected given the dataset. We can see that generally, higher
+incomes trend with a lower log-likelihood of being obese. The
+conclusions drawn from the activity levels also are in line with what we
+saw earlier on Figure~\ref{fig-activity}, though they are not at all
+what I expected as I believed exercise to have a negative correlation
+with obesity. I did not expect that the (Intercept) no
 
@@
 
-Figures \textbf{?@fig-page}, \textbf{?@fig-pgender}, and
-\textbf{?@fig-pedu} are recreations of the earlier graphs we saw
-showcasing the proportion of voters for Joe Biden by groups within each
-demographic, where the bar is the data from CCES 2020 while the point is
-the prediction generated by the model.
+Figures Figure~\ref{fig-pactivity} and Figure~\ref{fig-pobesity}, are
+recreations of the earlier graphs we saw showcasing the proportion of
+obesity by groups within categories, where the bar is the data from CDC
+while the point is the prediction generated by the model.
 
@@
 
-By observation, it is evident that the model's prediction was fairly
-close to what the CCES data showed. The trend remains the same for each
-result. However, gender and education were predicted to have a higher
-sentiment of voting for Joe Biden while age had a lower sentiment.
-
-It is worth nothing though that in Figure~\ref{fig-pactivity}, all age
-categories showed polling data of above 50\% to vote for Biden. Now with
-the lower average vote, the older age categories of ``45-64'' and
-``65+'' are leaning towards Trump. On Figure~\ref{fig-pobesity} for the
-lower education groups, the ``No-HS'' group which was at just about 50\%
-is now above it, and the ``High School Graduate'' is also closer to
-voting for Biden. While this may seem marginal, close elections where
-voter sentiment is close to 50\% would be largely impacted by changes
-like this. Although ultimately this is not an accurate predictor for who
-would win the election, only which groups would vote for who.
+We can see that the model prediction resembled the CDC data by trend,
+however for most of the categories within both
+Figure~\ref{fig-pactivity} and Figure~\ref{fig-pobesity}, the prediction
+for obesity was lower than the data suggested. Moderate cardio having
+such a high proportion of obesity as unexpected.
 
@@
 
+In this paper, we used real-world statistics on Americans' activity
+levels, income, and whether or not they were considered obese to find a
+correlation between these behavioural and socioeconomic factors and
+obesity. What we found was that contrary to popular belief, our model
+did not predict lower obesity levels for individuals who exercised more.
+In particular, there was a large proportion of individuals who did
+moderate cardio and were obese, but there was correlation with lower
+obesity and strength training. Generally, higher income levels were also
+associated with lower rates of obesity.
+
+While this data may not be indicative of reality, the lack of
+correlation between exercise and obesity may reveal truths about
+societal beliefs in terms of how to get fit. Like myself, many Americans
+grew up watching TV where commercial ad breaks would feature fad diets,
+fast 10 minute work outs, and buzz phrases like ``do this for 30 minutes
+a week to lose X lbs of fat!'' Fitness may be viewed with a dogmatic
+stigma due to media portrayals which push the idea of a `hardcore'
+lifestyle being required for a fit body.
+
+Due to this, it is possible that individuals believe in exercise being a
+greater factor for fat loss than nutrition. With the highest category of
+300 minutes of moderate cardio used in CDC's survey, an adult weighing
+160 lbs can expect to burn approximately 1825 calories in a week (Mayo
+Clinic Staff 2021). As a single pound of body fat contains roughly 3500
+calories, this would seem like half a pound of fat loss per week.
+However, this does not account for the possibility that an individual
+consumes more food with added physical activity, and 1825 calories
+spread out over a week is obly 260 calories per day. That equates to
+less than a small fries at McDonalds, or 2 or 3 eggs with oil.
+
+It is impossible to predict how an individual's body composition will
+change without knowing their nutritional information, specifically their
+caloric intake. There are various reasons why higher incomes might trend
+with lower obesity rates. People with higher incomes also tend to have
+achieved higher levels of education, which may be tied to access of
+information. They also have more access to help from professionals,
+better training facilities, and more food options. The study ``Nutrition
+quality of food purchases varies by household income: the SHoPPER
+study'' published in BMC Public Health highlights that lower-income
+households tend to purchase foods of lower nutritional quality compared
+to higher-income households. This is due to financial constraints
+limiting the purchase of healthier options like fruits and vegetables,
+leading to a higher purchase of less healthful foods such as frozen
+desserts or fast food. The study emphasizes that food purchasing
+patterns significantly mediate income differences in dietary intake
+quality (French et al. 2019).
+
 \hypertarget{weaknesses-and-next-steps}{%
@@
 
-\newpage
+As outlined previously, the data points were self reported which could
+lead to false data due to dishonesty or lack of care in measurement. The
+dataset was also filtered to remove incomplete datapoints, resulting in
+almost 90\% of the initial dataset being removed. The discrete
+categories used for both physical activity and income may oversimplify
+the spectrum of exercise habits which overlooks the nuances of
+individual physical activity patterns. The income cap at \$75,000 also
+fails to show variations in prevalence of obesity in higher income
+brackets.
+
+Future research should aim to incorporate objective measures of physical
+activity, perhaps through wearable technology, to diminish
+self-reporting biases. A broader dataset, possibly integrating direct
+measures of physical activity and detailed nutritional intake, would
+enable a more comprehensive analysis. Investigating the impact of higher
+income brackets beyond \$75,000 and accounting for regional
+cost-of-living differences could refine understanding of the
+income-obesity relationship. Additionally, longitudinal data could shed
+light on the temporal dynamics between exercise habits and BMI changes
+over time.
+
+Further, integrating geospatial analyses to assess environmental
+factors, such as access to recreational spaces and healthy food outlets,
+could add further context to obesity determinants. Given the
+multifaceted nature of obesity, interdisciplinary studies combining data
+from healthcare, urban planning, and social sciences could offer more
+holistic insights. \newpage
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-brfss}{}}%
-{``4 Comparison of Data Sources Used to Assess Obesity Prevalence and
-Trends.''} n.d. The National Academies of Sciences, Engineering,;
-Medicine; \url{https://nap.nationalacademies.org/read/23505/chapter/7}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-Modelsummary}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dataset}{}}%
-Centers for Disease Control and Prevention (CDC), National Center for
-Chronic Disease Prevention and Health Promotion, Division of Nutrition,
-Physical Activity, and Obesity. 2023. {``{Nutrition, Physical Activity,
-and Obesity - Behavioral Risk Factor Surveillance System}.''} Atlanta,
-Georgia:
+Centers for Disease Control and Prevention (CDC). 2023. {``{Nutrition,
+Physical Activity, and Obesity - Behavioral Risk Factor Surveillance
+System}.''} Atlanta, Georgia:
 \url{https://chronicdata.cdc.gov/Nutrition-Physical-Activity-and-Obesity/Nutrition-Physical-Activity-and-Obesity-Behavioral/hn4x-zwk7};
@@
 Journal of Preventive Medicine} 27: 154--62.
+
+\leavevmode\vadjust pre{\hypertarget{ref-nutritionquality}{}}%
+French, Simone A., Christy C. Tangney, Melissa M. Crane, Yamin Wang, and
+Bradley M. Appelhans. 2019. {``Nutrition Quality of Food Purchases
+Varies by Household Income: The SHoPPER Study.''} \emph{BMC Public
+Health} 19 (231).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-mayoclinic}{}}%
+Mayo Clinic Staff. 2021. {``Exercise for Weight Loss: Calories Burned in
+1 Hour.''}
+\url{https://www.mayoclinic.org/healthy-lifestyle/weight-loss/in-depth/exercise/art-20050999}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-chatgpt}{}}%
@@
 \url{https://openai.com/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cdcinitiatives}{}}%
-{``Overweight \& Obesity \textbar{} CDC.''} n.d. Centers for Disease
-Control; Prevention; \url{https://www.cdc.gov/obesity/index.html}.
 

### 2024-04-19 scripts/test.R

@@ -1,15 +1,67 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Tests cleaned data
+# Author: Victor Ma
+# Date: 19 April 2024
+# Contact: victo.ma@mail.utoronto.ca
+# In this script I use code from OpenAI's ChatGPT4. Chatlogs can be found in the usage.txt file.
 
 
-#### Workspace setup ####
+
+
 library(tidyverse)
-# [...UPDATE THIS...]
+# Check if files exist
+if (!file.exists("inputs/data/analysis_data/analysis_data.csv")) {
+  stop("CSV file does not exist.")
+}
 
-#### Test data ####
+if (!file.exists("inputs/data/analysis_data/analysis_data.parquet")) {
+  stop("Parquet file does not exist.")
+}
+
+# Load data
+csv_data <- read.csv("inputs/data/analysis_data/analysis_data.csv")
+parquet_data <- arrow::read_parquet("inputs/data/analysis_data/analysis_data.parquet")
+
+# Check columns
+required_columns <- c("obese", "activity", "income")
+if (!all(required_columns %in% names(csv_data))) {
+  stop("One or more required columns are missing in the CSV data.")
+}
+
+if (!all(required_columns %in% names(parquet_data))) {
+  stop("One or more required columns are missing in the Parquet data.")
+}
+
+# Check for duplicates
+if (any(duplicated(csv_data))) {
+  stop("There are duplicate rows in the CSV data.")
+}
+
+if (any(duplicated(parquet_data))) {
+  stop("There are duplicate rows in the Parquet data.")
+}
+
+# Validate data range for a hypothetical numeric column (if exists)
+# assuming 'BMI' is a numeric column for demonstration
+if (any(csv_data$BMI < 10 | csv_data$BMI > 50, na.rm = TRUE)) {
+  stop("BMI values out of expected range [10, 50] in CSV data.")
+}
+
+# Check factor levels
+expected_activity_levels <- c("No Activity", "Strength", "Cardio", "Cardio + Strength", "Double Cardio")
[…]

### 2024-04-19 paper/paper.qmd
fixed pdf issue

@@
 thanks: "Code and data are available at: https://github.com/bestmustard/activity-bmi"
-date: 03/16/2024
+date: 04/19/2024
 date-format: long
@@
 This includes tidyverse for data manipulation and visualization, knitr for report generation, modelsummary for model interpretation, and rstanarm for Bayesian regression modeling [@citeR; @Knitr; @Modelsummary; @Rstanarm; @Tidybayes; @Dplyr].
-Some portions including ggplot graphs and the "Data" and "Discussion" sections were written with the help of ChatGPT4 @chatgpt.
+Some portions including ggplot graphs and the "Data", "Summary Interpretation", and "Discussion" sections were written with the help of ChatGPT4 @chatgpt.
 
@@
 
-After creating the model, we can examine the summary in @fig-summary:
+We can see the model summary in  @tbl-modelresults:
 
-```{r out.height="20%"}
-#| label: fig-summary
-#| fig-cap: "Model summary"
+```{r}
 #| echo: false
+#| eval: true
 #| warning: false
 #| message: false
-obesity_likelihood <- readRDS(file ="../models/obesity_likelihood.rds")
 
-modelsummary(obesity_likelihood)
+library(rstanarm)
+
+first_model <-
+  readRDS(file = "../models/obesity_likelihood.rds")
 ```
 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory model of obesity based on activity and income level"
+#| warning: false
+
+modelsummary::modelsummary(
+  list(
+    "Obesity Likelihood" = first_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
 #### Summary Interpretation
@@
 
-(Intercept) (5.329): When all other variables are at their baseline levels (no physical activity and the lowest income category), the log odds of being obese are 5.329.
+(Intercept) (0.06): This is the log-odds of being obese when all other variables are zero.
+Since it's close to zero, it suggests that when no activity is accounted for and income is at its baseline (likely the lowest income category), the likelihood of being obese is near the 50% mark on the probability scale.
 
-activityCardio + Strength (-8.164): Engaging in both cardio and strength activities is associated with an 8.164 unit decrease in the log odds of being obese compared to the baseline of no activity.
+activityStrength (0.18): Engaging in strength training activities is associated with a 0.18 increase in the log-odds of being obese compared to no physical activity.
+The positive coefficient suggests a slight association with higher obesity odds, which might be unexpected.
 
-activityDouble Cardio (-4.518): Doing 300 minutes of moderate cardio or 150 minutes of vigorous cardio results in a 4.518 unit decrease in the log odds of being obese compared to the baseline.
+activityCardio (5.37): Engaging in cardio activities is associated with a substantial increase (5.37) in the log-odds of being obese.
+This result is counterintuitive as cardio exercises typically correlate with weight loss or lower obesity rates.
 
-activityNo Activity (-5.269): Having no leisure-time physical activity is associated with a 5.269 unit decrease in the log odds of being obese compared to the baseline.
-This negative value is unexpected, as we would typically predict no activity to increase the odds of obesity unless the baseline includes negative health behaviors even more associated with obesity.
+activityCardio + Strength (-2.90): Engaging in both cardio and strength activities is associated with a decrease in the log-odds of being obese.
+This suggests that combined exercise types might be effective at reducing the likelihood of obesity.
 
-activityStrength (-5.092): Engaging only in strength training activities is associated with a 5.092 unit decrease in the log odds of being obese compared to the baseline.
+activityDouble Cardio (0.75): Performing a double amount of cardio is associated with an increase in the log-odds of being obese.
+Like the cardio variable, this positive association is counter to typical expectations.
 
-income15-25 (0.096): Being in the \$15,000 to $25,000 income bracket is associated with a 0.096 unit increase in the log odds of being obese compared to the reference income category.
+For the income variables (reference being likely the lowest income category):
 
-income25-35 (-0.719): Being in the \$25,000 to $35,000 income bracket is associated with a 0.719 unit decrease in the log odds of being obese.
+income15-25 (0.10): Being in the $15,000 to $25,000 income bracket is associated with a slight increase in the log-odds of being obese.
 
-income35-50 (-0.990): Being in the \$35,000 to $50,000 income bracket is associated with a 0.990 unit decrease in the log odds of being obese.
+income25-35 (-0.72): This income bracket is associated with a decrease in the log-odds of being obese, suggesting that a higher income level correlates with lower obesity odds.
 
-income50-75 (-1.063): Being in the \$50,000 to $75,000 income bracket is associated with a 1.063 unit decrease in the log odds of being obese.
+income35-50 (-0.98): Similarly, this indicates a greater decrease in the log-odds of being obese, reinforcing the trend that higher income brackets correlate with lower obesity odds.
 
-income75+ (-0.401): Being in the over $75,000 income bracket is associated with a 0.401 unit decrease in the log odds of being obese.
+income50-75 (-1.06): This bracket sees an even further decrease in the log-odds of obesity.
+
+income75+ (-0.40): This suggests a smaller decrease in the log-odds of being obese for the highest income bracket compared to some of the lower brackets.
 
